---
title: "Jun To Do"
author: "D. Ford Hannum Jr."
date: "4/5/2021"
output: 
        html_document:
                toc: true
                toc_depth: 3
                code_folding: hide
                number_sections: false
                theme: united
                highlight: tango
---

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(Seurat)
library(ggplot2)
library(data.table)
library(MAST)
library(SingleR)
library(dplyr)
library(tidyr)
library(limma)
library(scRNAseq)
library(ggpubr)
library(xlsx)
library(ComplexHeatmap)
```

# LOADING DATA

```{r}
wbm <- readRDS('./data/v3/mpl.migr1.121420-2.rds')
table(wbm$Condition2)

wbm$figure.labels <- ifelse(wbm$v7.labels %in% c('MEP/MCP','MEP/ERP'),
                            ifelse(wbm$v7.labels == 'MEP/MCP', 'MkP','MEP'),
                            as.character(wbm$v7.labels))

wbm$figure.labels <- ifelse(wbm$figure.labels == 'Tcell', 'T-cell',
                            ifelse(wbm$figure.labels == 'Bcell','B-cell',
                                   ifelse(wbm$figure.labels == 'Bcell-Prog',
                                          'B-cell Prog.', wbm$figure.labels)))
wbm$UMAP1 <- wbm@reductions$umap@cell.embeddings[,1]
wbm$UMAP2 <- wbm@reductions$umap@cell.embeddings[,2]

wbm$figure.labels2 <- ifelse(wbm$figure.labels != 'MK',
                             as.character(wbm$figure.labels),
                            ifelse(wbm$UMAP1 < 0, 'MK-1','MK-2'))

wbm$figure.labels2 <- ifelse(wbm$figure.labels != 'Mono/Macro', 
                         as.character(wbm$figure.labels2),
                         ifelse(wbm$UMAP1 < -5, 'Mono/Macro-1','Mono/Macro-2'))

wbm$figure.labels <- factor(wbm$figure.labels,
                            levels = levels(as.factor(wbm$figure.labels))[c(5,3,1,2,10,9,4,6,8,7)])

Idents(wbm) <- wbm$figure.labels

# table(Idents(wbm))

mk1.cells <- rownames(wbm@meta.data[wbm$figure.labels == 'MK-1',])
mep1.cells <- rownames(wbm@meta.data[wbm$figure.labels == 'MkP',])

wbm$figure.labels3 <- ifelse(wbm$figure.labels2 %in% c('Mono/Macro-1','Mono/Macro-2'),
                             'Mono/Macro', as.character(wbm$figure.labels2))
color_pal <- c("#0072B2", "#CC79A7", "#009E73",'black',"#999999",
               "#E69F00", "#56B4E9","#D55E00", 'limegreen', 'violet')
color_pal2 <- c(color_pal, 'red')

wbm$Condition3 <- ifelse(wbm$Condition2 == 'Migr1', 'MigR1',
                         ifelse(wbm$Condition2 == 'enrMigr1', 'CD41+ enr. MigR1',
                                ifelse(wbm$Condition2 == 'Mpl', 'MPLW515L', 'CD41+ enr. MPLW515L')))

wbm$Condition3 <- factor(wbm$Condition3,
                         levels = levels(as.factor(wbm$Condition3))[c(3,4,1,2)])

wbm$Experiment <- ifelse(wbm$Condition2 %in% c('enrMigr1','Migr1'), 'MigR1','MPL')
```

```{r}
# wbm@meta.data
wbm$tsne1 <- wbm@reductions$tsne@cell.embeddings[,'tSNE_1']
wbm$tsne2 <- wbm@reductions$tsne@cell.embeddings[,'tSNE_2']

# write.table(wbm@meta.data, './data/v4/seurat.metadata.txt', sep = '\t', quote = F)
```


# 1 Dimension Reduction Plots 

```{r}
DimPlot(wbm, group.by = 'figure.labels3', cols = color_pal2, label = T, 
        repel = T) + NoLegend()
DimPlot(wbm, reduction = 'tsne', group.by = 'figure.labels3', cols = color_pal2,
        label = T, repel = T) + NoLegend()

DimPlot(wbm, reduction = 'pca', dims = c(1,2), group.by = 'figure.labels3', 
        cols = color_pal2,
        label = T, repel = T) + NoLegend() + ggtitle(NULL)

DimPlot(wbm, reduction = 'pca', dims = c(1,3), group.by = 'figure.labels3', 
        cols = color_pal2,
        label = T, repel = T) + NoLegend() + ggtitle(NULL)

DimPlot(wbm, reduction = 'pca', dims = c(1,4), group.by = 'figure.labels3', 
        cols = color_pal2,
        label = T, repel = T) + NoLegend() + ggtitle(NULL)

DimPlot(wbm, reduction = 'pca', dims = c(1,5), group.by = 'figure.labels3', 
        cols = color_pal2,
        label = T, repel = T) + NoLegend() + ggtitle(NULL)
```

# 2 Version2 CCA Batch Correction

**V2**

Checking to see if CCA batch correction is useful and how it may change our projection

```{r}
# int.list <- SplitObject(wbm, split.by = 'Condition3')
# 
# for (i in 1:length(int.list)){
#   print(i)
#   int.list[[i]] <- SCTransform(int.list[[i]], verbose = F)
# }
# 
# int.features <- SelectIntegrationFeatures(object.list = int.list, nfeatures = 3000)
# 
# options(future.globals.maxSize = 3145728000)
# 
# int.list <- PrepSCTIntegration(object.list = int.list, anchor.features = int.features,
#                                verbose = F)
# 
# int.anchors <- FindIntegrationAnchors(object.list = int.list, normalization.method = 'SCT',
#                                       anchor.features = int.features, verbose = F)
# 
# int.int <- IntegrateData(anchorset = int.anchors, normalization.method = 'SCT',
#                          verbose = F)
# 
# int <- int.int
# 
# int <- RunPCA(int, verbose = F)
# 
# int <- RunUMAP(int, dims = 1:30)
# 
# int <- FindNeighbors(int, dims = 1:30)
# 
# for (i in seq(0,1,by = .1)){
#   temp <- FindClusters(int, resolution = i, verbose = F)
#   print(DimPlot(temp, label =T , repel= T))
# }
# 
# int <- FindClusters(int, resolution = 0.3, verbose = F)
# 
# DimPlot(int)
# 
# DimPlot(int, label = T, repel = T)
# 
# saveRDS(int, './data/v4/cca.integrated.rds')

int <- readRDS('./data/v4/cca.integrated.rds')

for (i in seq(0,1,by = .1)){
  temp <- FindClusters(int, resolution = i, verbose = F)
  print(DimPlot(temp, label =T , repel= T))
}
```

## 2a Cluster Count Tables 

```{r}
DimPlot(wbm, group.by = 'figure.labels3', label = T , repel = T)

df <- as.data.frame(table(int$figure.labels3, int$seurat_clusters))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','Mono/Macro','MkP','CMP','MEP','MK-2',
            'Erythroid','T-cell','B-cell Prog.')

df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

```{r}
int$int.label <- factor(int$seurat_clusters,
                        labels = c('Gran','Gran','Gran','B-cell','Gran','Mono/Macro',
                        'MkP','CMP','MEP/B-cell','Mono/Macro/MK-2','Erythroid','MkP',
                        'T-cell','B-cell','B-cell','B-cell Prog.'))
# table(int$int.label, int$seurat_clusters)

DimPlot(int, group.by = 'int.label', label = T, repel = T, cols = color_pal)
FeaturePlot(int, features = c('Itga2b','Pf4'))
```


### Experiment Count Tables

```{r}
DimPlot(int, group.by = 'Condition2')
DimPlot(int, split.by = 'Condition2', ncol = 2) + theme_bw()
```

```{r}
df <- as.data.frame(table(int$Condition2, int$seurat_clusters))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

# order_ <- c('MK-1','Granulocyte','B-cell','Mono/Macro','MkP','CMP','MEP','MK-2',
#             'Erythroid','T-cell','B-cell Prog.')
# 
# df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text(size = 4) + xlab('Experiment') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .15)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + 
  geom_text(size = 4) + xlab('Experiment') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

### Variable Features

```{r}
hv.genes <- VariableFeatures(int)
hv.genes.org <- VariableFeatures(wbm)

length(intersect(hv.genes, hv.genes.org))
```

Around half the highly variable genes overlap

## 2a.2 Looking at another resolution

Perhaps we were looking at too low of a resolution with the integrated model.

```{r}
int <- FindClusters(int,resolution = 1, verbose = F)

Idents(int) <- int$seurat_clusters

DimPlot(int, label = T, repel = T)
```

```{r}
# DimPlot(wbm, group.by = 'figure.labels3', label = T , repel = T)

df <- as.data.frame(table(int$figure.labels3, int$seurat_clusters))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','Mono/Macro','MkP','CMP','MEP','MK-2',
            'Erythroid','T-cell','B-cell Prog.')

df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

#### int.labels

```{r}
int$int.label <- factor(int$seurat_clusters,
                        labels = c('Gran','Gran','Gran','B-cell','Gran','Gran','Gran','MkP',
                                   'Mono/Macro', 'Gran','CMP','B-cell','Gran','Gran',
                                   'Mono/Macro','Erythroid','B-cell','MEP',
                                   'T-cell','MkP','Gran', 'B-cell',
                                   'B-cell','Mono/Macro','B-cell Prog.', 'MK-2'))

# saveRDS(int, './data/v4/cca.integrated.rds')

table(int$int.label, int$seurat_clusters)

DimPlot(int, group.by = 'int.label', label = T, repel = T)
FeaturePlot(int, features = c('Itga2b','Pf4'))
```

#### Further cross-tab with new naming

```{r}
df <- as.data.frame(table(int$figure.labels3, int$int.label))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','MkP','Mono/Macro','CMP','Erythroid',
            'MEP',
            'T-cell','B-cell Prog.','MK-2')

df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

### Experiment Count Tables

```{r}
DimPlot(int, group.by = 'Condition2')
DimPlot(int, split.by = 'Condition2', ncol = 2) + theme_bw()
```

```{r}
df <- as.data.frame(table(int$Condition2, int$int.label))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

# order_ <- c('MK-1','Granulocyte','B-cell','Mono/Macro','MkP','CMP','MEP','MK-2',
#             'Erythroid','T-cell','B-cell Prog.')
# 
# df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text(size = 4) + xlab('Experiment') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .15)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + 
  geom_text(size = 4) + xlab('Experiment') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

## 2b Combining w/ Original Data

```{r}
df2 <- as.data.frame(table(wbm$Condition2, wbm$figure.labels3))
df2$var1.total <- NA

for (i in unique(df2$Var1)){
  print(i)
  df2[df2$Var1 == i,]$var1.total <- sum(df2[df2$Var1 == i,]$Freq)
}

df2$var2.total <- NA
for (i in unique(df2$Var2)){
  print(i)
  df2[df2$Var2 ==i,]$var2.total <- sum(df2[df2$Var2 ==i,]$Freq)
}

df2$perc.var1 <- df2$Freq / df2$var1.total
df2$perc.var2 <- df2$Freq / df2$var2.total

ggplot(df2, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .15)

ggplot(df2, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + 
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

```{r}
df$status <- 'post-correction'
df2$status <- 'pre-correction'
df2$Var2 <- ifelse(df2$Var2 == 'Granulocyte', 'Gran', as.character(df2$Var2))
dfc <- rbind(df2,df)

table(dfc$Var1, dfc$Var2)

dfc$status <- factor(dfc$status, levels = c('pre-correction','post-correction'))

for (exp in unique(dfc$Var1)){
  # print(exp)
  tmp <- dfc[dfc$Var1 == exp,]
  print(ggplot(tmp, aes(x = Var2, y = perc.var1, fill = status, label = Freq)) +
    geom_bar(position = 'dodge', stat = 'identity') + 
    ggtitle (exp) +
    theme_bw() +
    geom_text(position = position_dodge(width = .9), vjust = -.25, size = 2.5) + 
    theme(axis.text.x = element_text(angle = 45, hjust =1))
  )
}

```

### Two Tile Charts in One

```{r}
df2 <- as.data.frame(table(wbm$Condition2, wbm$figure.labels3))
df2$var1.total <- NA

for (i in unique(df2$Var1)){
  print(i)
  df2[df2$Var1 == i,]$var1.total <- sum(df2[df2$Var1 == i,]$Freq)
}

df2$var2.total <- NA
for (i in unique(df2$Var2)){
  print(i)
  df2[df2$Var2 ==i,]$var2.total <- sum(df2[df2$Var2 ==i,]$Freq)
}

df2$perc.var1 <- df2$Freq / df2$var1.total
df2$perc.var2 <- df2$Freq / df2$var2.total

df <- as.data.frame(table(int$Condition2, int$int.label))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

df$var1.plus <- paste0('2_',df$Var1)
df2$var1.plus <- paste0('1_',df2$Var1)

df2$Var2 <- ifelse(df2$Var2 == 'Granulocyte', 'Gran',as.character(df2$Var2))

df3 <- rbind(df,df2)



ggplot(df3, aes(x = var1.plus, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .15)

ggplot(df3, aes(x = var1.plus, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + 
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

### Plots

```{r}
DimPlot(int, group.by = 'int.label', label = T, repel = T) + ggtitle('V2 Clustering/V2 Naming')
DimPlot(int, group.by = 'figure.labels3', label = T, repel = T) + 
  ggtitle('V2 Clustering/V2 Naming')

int$clustering4.labels <- v4$int.labels

DimPlot(int, group.by = 'clustering4.labels', label = T, repel = T) + 
  ggtitle('V2 Clustering/V4 Naming')
```


### Variable Features

```{r}
hv.genes <- VariableFeatures(int)
hv.genes.org <- VariableFeatures(wbm)

length(intersect(hv.genes, hv.genes.org))
```

Going with the second clustering!

## 2c Cluster Centroids

### All Genes

```{r}
Idents(int) <- int$int.label

av <- AverageExpression(int)$RNA

pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(av, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

ggplot(tc, aes(x = ct1, y = ct2, fill = Correlation)) + geom_tile() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### HV Genes

```{r}
pcs <- prcomp(av[hv.genes,])
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(av[hv.genes,], pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

ggplot(tc, aes(x = ct1, y = ct2, fill = Correlation)) + geom_tile() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## 2d Cluster Centroids by Sample

### By Experiment

#### All Genes

```{r}
Idents(int) <- paste0(int$int.label, '_', int$Condition2)

av <- AverageExpression(int)$RNA

pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(av, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

# order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$exp,
        column_split = order.rows$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

#### HV Genes

```{r}
av <- av[hv.genes,]
pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(av, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

# order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$exp,
        column_split = order.rows$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

### By Cell Type

#### All Genes

```{r}
# Idents(int) <- paste0(int$int.label, '_', int$Condition2)

av <- AverageExpression(int)$RNA

pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(av, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$ct, order.rows$exp),]$coln
order.rows <- order.rows[order(order.rows$pc,order.rows$ct, order.rows$exp),]
# order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$ct,
        column_split = order.rows$ct,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

#### HV Genes

```{r}
av <- av[hv.genes,]
pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(av, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

# order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$ct,
        column_split = order.rows$ct,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

### With Column V5 By Cell Type

Adding a column for post-corrected cluster centroids

#### All Genes

```{r}
Idents(int) <- paste0(int$int.label, '_', int$Condition2)

av <- AverageExpression(int)$RNA
head(av)

Idents(int) <- int$int.label
av2 <- AverageExpression(int)
av2 <- av2$integrated

pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)
colnames(av2) <- paste0(colnames(av2),'_Int')

dim(av)

av.c <- cbind(av, pc.vectors[,1:10])

av.c2 <- merge(av2, av.c, by = 0, all.y = T)

rownames(av.c2) <- av.c2$Row.names

av.c2 <- av.c2[,-1]

temp.cor <- cor(av.c2, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])
order.rows$int <- ifelse(order.rows$exp == 'Int', 1,0)

order.rows.order <- order.rows[order(order.rows$pc,order.rows$ct,
                                     order.rows$int, order.rows$exp),]$coln

order.rows <- order.rows[order(order.rows$pc,order.rows$ct,
                               order.rows$int,order.rows$exp),]
# order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred','Int' = 'black','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$ct,
        column_split = order.rows$ct,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

#### HV Genes

```{r}
av <- av[hv.genes,]
pcs <- prcomp(av)
pc.vectors <- pcs$x

Idents(int) <- int$int.label
av2 <- AverageExpression(int)
av2 <- av2$integrated

colnames(av2) <- paste0(colnames(av2),'_Int')

dim(av)

av.c <- cbind(av, pc.vectors[,1:10])

av.c2 <- merge(av2, av.c, by = 0, all.y = T)

rownames(av.c2) <- av.c2$Row.names

av.c2 <- av.c2[,-1]

temp.cor <- cor(av.c2, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])
order.rows$int <- ifelse(order.rows$exp == 'Int', 1,0)

order.rows.order <- order.rows[order(order.rows$pc,order.rows$ct,
                                     order.rows$int, order.rows$exp),]$coln

order.rows <- order.rows[order(order.rows$pc,order.rows$ct,
                               order.rows$int,order.rows$exp),]
# order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred','Int' = 'black','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$ct,
        column_split = order.rows$ct,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```


## 2e Cluster Centroid Deltas

### Using Uncorrected Values

Using top 10 PCs from cluster centroids

```{r}
Idents(int) <- paste0(int$int.label, '_', int$Condition2)
av <- AverageExpression(int)$RNA
# head(av)

av <- as.data.frame(t(av))

av$exp <- tstrsplit(rownames(av), '_', keep = 2)[[1]]

av$ct <- tstrsplit(rownames(av), '_', keep = 1)[[1]]

outs <- split(av, f = av$ct)
genes <- colnames(outs[[1]])[!colnames(outs[[1]]) %in% c('ct','exp')]
out.df <- data.frame(na.row = rep(NA,length(genes)),
                     row.names = genes)

for (p in outs){
  temp.df <- t(p[,!(colnames(p) %in% c('ct','exp'))])
  ct <- tstrsplit(colnames(temp.df)[1], '_', keep = 1)[[1]]
  colnames(temp.df) <- tstrsplit(colnames(temp.df), '_', keep = 2)[[1]]
  
  combos <- combn(1:length(colnames(temp.df)), 2)
  
  if (dim(temp.df)[2] == 4){
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1','enrMpl')]
  }
  else{
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1')]
  }
  
  for (i in 1:dim(combos)[2]){
    clm1 <- combos[1,i]
    clm2 <- combos[2,i]
    # print(paste0(clm1, clm2))
    
    if (i == 1){
      new.df.temp <- data.frame(V1 = temp.df[,clm1] - temp.df[,clm2])
      
    }
    else{
      new.df.temp <- cbind(new.df.temp, temp.df[,clm1] - temp.df[,clm2])
    }
    colnames(new.df.temp)[i] <- paste0(colnames(temp.df)[clm1],
                                       '_vs_', colnames(temp.df)[clm2],
                                       '.',ct)
  }
  
  out.df <- cbind(out.df, new.df.temp)
  
}

# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 2)[[1]]))
# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 1)[[1]]))

out.df <- out.df[,-1]
```

#### By Cluster

##### All Genes

```{r}
# head(out.df)
Idents(int) <- int$int.label
av <- AverageExpression(int)$RNA

pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(out.df, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue', 'Migr1_vs_enrMigr1' = 'red', 
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred', 'Mpl_vs_enrMpl' = 'orange',
                   'enrMigr1_vs_enrMpl' = 'pink','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$ct,
        column_split = order.rows$ct,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

##### HV Genes

```{r}
# head(out.df)
# Idents(int) <- int$int.label
# av <- AverageExpression(int)$RNA

pcs <- prcomp(av[hv.genes,])
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(out.df[hv.genes,], pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue', 'Migr1_vs_enrMigr1' = 'red', 
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred', 'Mpl_vs_enrMpl' = 'orange',
                   'enrMigr1_vs_enrMpl' = 'pink','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$ct,
        column_split = order.rows$ct,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

#### By Experiment

##### All Genes

```{r}
# head(out.df)
Idents(int) <- int$int.label
av <- AverageExpression(int)$RNA

pcs <- prcomp(av)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(out.df, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$ct, order.rows$exp),]$coln

order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue', 'Migr1_vs_enrMigr1' = 'red', 
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred', 'Mpl_vs_enrMpl' = 'orange',
                   'enrMigr1_vs_enrMpl' = 'pink','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$exp,
        column_split = order.rows$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

##### HV Genes

```{r}
# head(out.df)
# Idents(int) <- int$int.label
# av <- AverageExpression(int)$RNA

pcs <- prcomp(av[hv.genes,])
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(out.df[hv.genes,], pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$ct, order.rows$exp),]$coln

order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue', 'Migr1_vs_enrMigr1' = 'red', 
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred', 'Mpl_vs_enrMpl' = 'orange',
                   'enrMigr1_vs_enrMpl' = 'pink','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$exp,
        column_split = order.rows$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

### Post-correct

Using top 10 PCs from cluster centroids

```{r}
Idents(int) <- paste0(int$int.label, '_', int$Condition2)
av <- AverageExpression(int)$integrated

# head(av)

av <- as.data.frame(t(av))

av$exp <- tstrsplit(rownames(av), '_', keep = 2)[[1]]

av$ct <- tstrsplit(rownames(av), '_', keep = 1)[[1]]

outs <- split(av, f = av$ct)
genes <- colnames(outs[[1]])[!colnames(outs[[1]]) %in% c('ct','exp')]
out.df <- data.frame(na.row = rep(NA,length(genes)),
                     row.names = genes)

for (p in outs){
  temp.df <- t(p[,!(colnames(p) %in% c('ct','exp'))])
  ct <- tstrsplit(colnames(temp.df)[1], '_', keep = 1)[[1]]
  colnames(temp.df) <- tstrsplit(colnames(temp.df), '_', keep = 2)[[1]]
  
  combos <- combn(1:length(colnames(temp.df)), 2)
  
  if (dim(temp.df)[2] == 4){
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1','enrMpl')]
  }
  else{
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1')]
  }
  
  for (i in 1:dim(combos)[2]){
    clm1 <- combos[1,i]
    clm2 <- combos[2,i]
    # print(paste0(clm1, clm2))
    
    if (i == 1){
      new.df.temp <- data.frame(V1 = temp.df[,clm1] - temp.df[,clm2])
      
    }
    else{
      new.df.temp <- cbind(new.df.temp, temp.df[,clm1] - temp.df[,clm2])
    }
    colnames(new.df.temp)[i] <- paste0(colnames(temp.df)[clm1],
                                       '_vs_', colnames(temp.df)[clm2],
                                       '.',ct)
  }
  
  out.df <- cbind(out.df, new.df.temp)
  
}

# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 2)[[1]]))
# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 1)[[1]]))

out.df <- out.df[,-1]
```

#### By Cluster

##### HV Genes

```{r}
Idents(int) <- int$int.label
av2 <- AverageExpression(int)$integrated

pcs <- prcomp(av2)
pc.vectors <- pcs$x
# head(pc.vectors)

av.c <- cbind(out.df, pc.vectors[,1:10])

temp.cor <- cor(av.c, use = 'pairwise.complete.obs', method = 'spearman')

tc <- as.data.frame(temp.cor)

tc <- gather(tc, key = 'ct1', value = 'Correlation', factor_key = T)
tc$ct2 <- rep(unique(tc$ct1), dim(tc)[1]/length(unique(tc$ct1)))

order.rows <- data.frame(coln = colnames(temp.cor))
# order.rows
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows.order <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue', 'Migr1_vs_enrMigr1' = 'red', 
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred', 'Mpl_vs_enrMpl' = 'orange',
                   'enrMigr1_vs_enrMpl' = 'pink','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$ct,
        column_split = order.rows$ct,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

#### By Experiment

##### HV Genes

```{r}
order.rows.order <- order.rows[order(order.rows$pc,order.rows$ct, order.rows$exp),]$coln

# order.rows

col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Gran' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog' = '#CC79A7',
                        'MkP' = 'violet','MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue', 'Migr1_vs_enrMigr1' = 'red', 
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred', 'Mpl_vs_enrMpl' = 'orange',
                   'enrMigr1_vs_enrMpl' = 'pink','PC' = 'black'))

temp.cor2 <- temp.cor[order.rows.order, order.rows.order]


column_ha = HeatmapAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col)

row_ha = rowAnnotation(CellType = order.rows$ct,Exp = order.rows$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = order.rows$exp,
        column_split = order.rows$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

# *VERSION 3*

Doing the integration across all genes. This could've been accomplished with the first integration but needed to add another tag to the IntegrateData command. For the most part just reusing the same code as what was used in #2 (i.e. V2)

```{r}
# int.list <- SplitObject(wbm, split.by = 'Condition3')
# 
# for (i in 1:length(int.list)){
#   print(i)
#   int.list[[i]] <- SCTransform(int.list[[i]], verbose = F)
# }
# 
# int.features <- SelectIntegrationFeatures(object.list = int.list, 
#                                           nfeatures = 3000)
# 
# options(future.globals.maxSize = 3145728000)
# 
# int.list <- PrepSCTIntegration(object.list = int.list, anchor.features = int.features,
#                                verbose = F)
# 
# int.anchors <- FindIntegrationAnchors(object.list = int.list, 
#                                       normalization.method = 'SCT',
#                                       anchor.features = int.features, verbose = F)
# 
# all.genes <- rownames(wbm)
# 
# int.int <- IntegrateData(anchorset = int.anchors, normalization.method = 'SCT',
#                          verbose = F, features.to.integrate = all.genes)
# 
# int <- int.int
# 
# int <- RunPCA(int, verbose = F)
# 
# int <- RunUMAP(int, dims = 1:30)
# 
# int <- RunTSNE(int, dims = 1:30)
# 
# int <- FindNeighbors(int, dims = 1:30)
# 
# for (i in seq(0,1,by = .1)){
#   temp <- FindClusters(int, resolution = i, verbose = F)
#   print(DimPlot(temp, label =T , repel= T))
# }
# 
# int <- FindClusters(int, resolution = 1, verbose = F)
# 
# DimPlot(int)
# 
# DimPlot(int, label = T, repel = T)
# 
# saveRDS(int, './data/v4/cca.integrated.all.rds')

int2 <- readRDS('./data/v4/cca.integrated.all.rds')
```

## New Clustering 2 w/ more Centroids

```{r}
x <- int2@assays$integrated@scale.data
dim(x)

clusters <- kmeans(t(x), centers = 26, iter.max = 20)
int2$all.int.clusters2 <- clusters$cluster

table(int2$figure.labels3, int2$all.int.clusters2)
```

## Cluster Distribution 


```{r}
df <- as.data.frame(table(int2$figure.labels3, int2$all.int.clusters2))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','Mono/Macro','MkP','CMP','MEP','MK-2',
            'Erythroid','T-cell','B-cell Prog.')

df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

```{r}
new.levels <- c('Gran','B-cell/MkP','B-cell','MkP','MkP','B-cell/CMP','Gran',
                'MK-2','MkP','Gran','Erythroid','Gran','Gran','Mono/Macro','Mono/Macro',
                'Gran','MEP','Gran','Mono/Macro','CMP','MEP','T-cell','B-cell',
                'Gran','B-cell','B-cell Prog.')

int2$all.int.clusters2.label <- factor(int2$all.int.clusters2,
                                      labels = new.levels)

table(int2$all.int.clusters2, int2$all.int.clusters2.label)

table(int2$all.int.clusters2.label, int2$figure.labels3)

saveRDS(int2,'./data/v4/cca.integrated.all.rds')

# int2 <- readRDS('./data/v4/cca.integrated.all.rds')
```


## New Tile w/ 3 Clusterings

```{r}

# Version 1

df2 <- as.data.frame(table(wbm$Condition2, wbm$figure.labels3))
df2$var1.total <- NA

for (i in unique(df2$Var1)){
  print(i)
  df2[df2$Var1 == i,]$var1.total <- sum(df2[df2$Var1 == i,]$Freq)
}

df2$var2.total <- NA
for (i in unique(df2$Var2)){
  print(i)
  df2[df2$Var2 ==i,]$var2.total <- sum(df2[df2$Var2 ==i,]$Freq)
}

df2$perc.var1 <- df2$Freq / df2$var1.total
df2$perc.var2 <- df2$Freq / df2$var2.total

# Version 2

df <- as.data.frame(table(int$Condition2, int$int.label))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# Version 3

df4 <- as.data.frame(table(int2$Condition2, int2$all.int.clusters2.label))
df4$var1.total <- NA

for (i in unique(df4$Var1)){
  print(i)
  df4[df4$Var1 == i,]$var1.total <- sum(df4[df4$Var1 == i,]$Freq)
}

df4$var2.total <- NA
for (i in unique(df4$Var2)){
  print(i)
  df4[df4$Var2 ==i,]$var2.total <- sum(df4[df4$Var2 ==i,]$Freq)
}

df4$perc.var1 <- df4$Freq / df4$var1.total
df4$perc.var2 <- df4$Freq / df4$var2.total

df4$var1.plus <- paste0('3_',df4$Var1)
df$var1.plus <- paste0('2_',df$Var1)
df2$var1.plus <- paste0('1_',df2$Var1)

df2$Var2 <- ifelse(df2$Var2 == 'Granulocyte', 'Gran',as.character(df2$Var2))

df3 <- rbind(df2,rbind(df,df4))



ggplot(df3, aes(x = var1.plus, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .15)

ggplot(df3, aes(x = var1.plus, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + 
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```


## Cluster Centroids

## Cluster Deltas

# *VERSION 4*

Instead of doing kmeans as in Version 3, we are using the all integrated genes across cells as input into Seurat and running everything again.

```{r}
int2@assays$integrated@data[1:5,1:5] 
dim(int2@assays$integrated@data)

v4 <- CreateSeuratObject(counts = int2@assays$integrated@data, 
                         assay = 'integrated', 
                         project = 'Version4')

v4 <- NormalizeData(v4, normalization.method = 'LogNormalize', 
                    scale.factor = 10000)

v4 <- ScaleData(v4, features = rownames(v4))

v4 <- FindVariableFeatures(v4, selection.method = 'vst', nfeatures = 2000)

v4 <- SetAssayData(v4, assay = 'integrated', slot = "scale.data", 
                   new.data = as.matrix(int2@assays$integrated@data))

v4 <- RunPCA(v4, features = VariableFeatures(v4),
             assay = 'integrated')

v4 <- FindNeighbors(v4, dims = 1:20)
v4 <- FindClusters(v4, resolution = 1)

v4 <- RunUMAP(v4, dims = 1:20)

DimPlot(v4)
```

## Cross Tabulation

### With Cluster Numbers

```{r}
v4$figure.labels3 <- wbm$figure.labels3

df <- as.data.frame(table(v4$figure.labels3, v4$seurat_clusters))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','Mono/Macro','MkP','CMP','MEP','MK-2',
            'Erythroid','T-cell','B-cell Prog.')

df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

```{r}
new.levels <- c('Gran','Gran','B-cell',rep('Gran',4), 'MkP','Mono/Macro','CMP',
                'Gran','B-cell',rep('Gran',3),'Mono/Macro','Erythroid','B-cell','MEP',
                'T-cell','MkP','B-cell','B-cell','Mono/Macro','B-cell Prog.')

v4$int.labels <- factor(v4$seurat_clusters,
                                      labels = new.levels)

table(v4$int.labels, v4$seurat_clusters)

table(v4$int.labels, v4$figure.labels3)
```

### W Cluster Names

```{r}
df <- as.data.frame(table(v4$figure.labels3, v4$int.labels))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','MkP','Mono/Macro','CMP','Erythroid',
            'MEP','T-cell','B-cell Prog.','MK-2')

df$Var1 <- factor(df$Var1, levels = order_)


ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

## Projections

```{r}
DimPlot(v4, group.by = 'int.labels', label = T, repel = T) + ggtitle('V4 Cluster/V4 Labels')
DimPlot(v4, group.by = 'figure.labels3', label = T, repel = T) + ggtitle('V4 Cluster/V1 Labels')

v4$v3.labels <- int2$all.int.clusters2.label

DimPlot(v4, group.by = 'v3.labels', label = T, repel = T) + ggtitle('V4 Cluster/V3 Labels')

v4$v2.labels <- int$int.label

DimPlot(v4, group.by = 'v2.labels', label = T, repel = T) + ggtitle('V4 Cluster/V3 Labels')

```


# *VERSION 1*

## Version Comparisons

### By Experiment

```{r}

# Version 1

df2 <- as.data.frame(table(wbm$Condition2, wbm$figure.labels3))
df2$var1.total <- NA

for (i in unique(df2$Var1)){
  print(i)
  df2[df2$Var1 == i,]$var1.total <- sum(df2[df2$Var1 == i,]$Freq)
}

df2$var2.total <- NA
for (i in unique(df2$Var2)){
  print(i)
  df2[df2$Var2 ==i,]$var2.total <- sum(df2[df2$Var2 ==i,]$Freq)
}

df2$perc.var1 <- df2$Freq / df2$var1.total
df2$perc.var2 <- df2$Freq / df2$var2.total

# Version 2

df <- as.data.frame(table(int$Condition2, int$int.label))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# Version 3
v4$Condition2 <- wbm$Condition2
df4 <- as.data.frame(table(v4$Condition2, v4$int.labels))
df4$var1.total <- NA

for (i in unique(df4$Var1)){
  print(i)
  df4[df4$Var1 == i,]$var1.total <- sum(df4[df4$Var1 == i,]$Freq)
}

df4$var2.total <- NA
for (i in unique(df4$Var2)){
  print(i)
  df4[df4$Var2 ==i,]$var2.total <- sum(df4[df4$Var2 ==i,]$Freq)
}

df4$perc.var1 <- df4$Freq / df4$var1.total
df4$perc.var2 <- df4$Freq / df4$var2.total

df4$var1.plus <- paste0('4_',df4$Var1)
df$var1.plus <- paste0('2_',df$Var1)
df2$var1.plus <- paste0('1_',df2$Var1)

df2$Var2 <- ifelse(df2$Var2 == 'Granulocyte', 'Gran',as.character(df2$Var2))

df3 <- rbind(df2,rbind(df,df4))



ggplot(df3, aes(x = var1.plus, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .15)

ggplot(df3, aes(x = var1.plus, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + 
  geom_text(size = 4) + xlab('Experiment') + ylab ('Original Clusters') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

```

### Cross Tabulation

#### v1 Y-axis

```{r}
df <- as.data.frame(table(v4$figure.labels3, v4$int.labels))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','MkP','Mono/Macro','CMP','Erythroid',
            'MEP','T-cell','B-cell Prog.','MK-2')

df$Var1 <- factor(df$Var1, levels = order_)

df2 <- df

df <- as.data.frame(table(int$figure.labels3, int$int.label))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

# df

order_ <- c('MK-1','Granulocyte','B-cell','MkP','Mono/Macro','CMP','Erythroid',
            'MEP',
            'T-cell','B-cell Prog.','MK-2')

df$Var1 <- factor(df$Var1, levels = order_)

df2$Var2 <- paste0('4_', df2$Var2)
df$Var2 <- paste0('2_', df$Var2)

df <- rbind(df,df2)

order_ <- c('B-cell','B-cell Prog.', 'CMP','Erythroid','Granulocyte','MK-1','MEP','MK-2','MkP',
            'Mono/Macro','T-cell')

df$Var1 <- factor(df$Var1, levels = order_)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(size = 4) + xlab('Original Clusters') + ylab ('CCA Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

#### v2 Y-axis

```{r}
df <- as.data.frame(table(v4$v2.labels, v4$int.labels))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

df2 <- df

wbm$version2.labels <- int$int.label

df <- as.data.frame(table(wbm$version2.labels, wbm$figure.labels3))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

df2$Var2 <- paste0('4_', df2$Var2)
df$Var2 <- paste0('1_', df$Var2)

df <- rbind(df,df2)

order_ <- c('B-cell','B-cell Prog.', 'CMP','Erythroid','Gran','MK-1','MEP','MK-2','MkP',
            'Mono/Macro','T-cell')

df$Var1 <- factor(df$Var1, levels = order_)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(size = 4) + xlab('v2 Clusters') + ylab ('Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(size = 4) + xlab('v2 Clusters') + ylab ('Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

#### v4 Y-axis

```{r}
int$version4.label <- v4$int.labels
df <- as.data.frame(table(int$version4.label, int$int.label))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

df2 <- df

wbm$version4.labels <- v4$int.labels

df <- as.data.frame(table(wbm$version4.labels, wbm$figure.labels3))
df$var1.total <- NA

for (i in unique(df$Var1)){
  print(i)
  df[df$Var1 == i,]$var1.total <- sum(df[df$Var1 == i,]$Freq)
}

df$var2.total <- NA
for (i in unique(df$Var2)){
  print(i)
  df[df$Var2 ==i,]$var2.total <- sum(df[df$Var2 ==i,]$Freq)
}

df$perc.var1 <- df$Freq / df$var1.total
df$perc.var2 <- df$Freq / df$var2.total

df2$Var2 <- paste0('4_', df2$Var2)
df$Var2 <- paste0('1_', df$Var2)

df <- rbind(df,df2)

order_ <- c('B-cell','B-cell Prog.', 'CMP','Erythroid','Gran','MK-1','MEP','MK-2','MkP',
            'Mono/Macro','T-cell')

df$Var1 <- factor(df$Var1, levels = order_)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var1, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(size = 4) + xlab('v4 Clusters') + ylab ('Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(df, aes(x = Var1, y = Var2, fill = perc.var2, label = Freq)) +
  geom_tile() + theme_bw() + coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(size = 4) + xlab('v4 Clusters') + ylab ('Clusters') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

### Projections

```{r}
DimPlot(wbm, group.by = 'figure.labels3', label = T, repel = T) + 
  ggtitle('V1 Clustering/V1 Naming')

DimPlot(wbm, group.by = 'version2.labels', label = T, repel = T) + 
  ggtitle('V1 Clustering/V2 Naming')

DimPlot(wbm, group.by = 'version4.labels', label = T, repel = T) + 
  ggtitle('V1 Clustering/V4 Naming')
```


# 3 Cluster Centroid Batch Effects

Getting the cluster centroid for each cluster by experiment (11x4, 44), then taking deltas between batch columns (11x6, 66 in total). 

i.e. MigR1 Gran minus MPL Gran would be the first column then MigR1 B-cell minus MPL B-cell. So this would all be comparing group 1 (MigR1) vs group 2 (MPL) across all clusters. Then it would be group 1 (MigR1) vs group 2 (enrMigR1), and so on for all six combinations.

It can be either organize be comparison between groups. But then also organize by clusters, to see how things are different between the clusters.

## 3a Cluster by Sample 

```{r}
wbm$ident.combo <- paste0(wbm$Condition, '_' ,wbm$figure.labels3)

# table(wbm$ident.combo)

# table(wbm$Condition3, wbm$figure.labels3)

# DimPlot(wbm, group.by = 'ident.combo')

Idents(wbm) <- wbm$ident.combo

av <- AverageExpression(wbm)

av <- av$RNA

# head(av)

key_ <- data.frame(ident.combo = colnames(av))

rownames(key_) <- 1:length(colnames(av))

key_$exp <- tstrsplit(key_$ident.combo, '_', keep = 1)[[1]]
key_$cluster <- tstrsplit(key_$ident.combo, '_', keep = 2)[[1]]

# temp.cor

# summary(key_$ident.combo == rownames(temp.cor))

```

### PC Plots

```{r}
pca <- prcomp(av)

pc.vectors <- pca$x

pca.sd <- pca$sdev
pca <- as.data.frame(pca$rotation)

pca.sd <- data.frame(pc = 1:length(pca.sd),
                     sd = pca.sd)

ggplot(pca.sd[1:20,], aes(x = pc, y = sd)) + geom_point() + geom_line() + theme_bw()
# colnames(pca) <- paste0(colnames(pca),' (',round(pca.sd,2),')')

pca$exp <- tstrsplit(rownames(pca),'_', keep = 1)[[1]]
pca$cluster <- tstrsplit(rownames(pca),'_', keep = 2)[[1]]
# pca

combos <- combn(paste0('PC',1:5),2)
# combos

# colnames(pca)[1:5] <- paste0(colnames(pca)[1:5], ' (SD = ', pca.sd[1:5], ')')

for (i in 1:dim(combos)[2]){
  # print(combos[1,i])
  temp.x <- combos[1,i]
  # print(combos[2,i])
  temp.y <- combos[2,i]
  
  print(ggplot(pca, aes_string(x = temp.x, y = temp.y, color = 'cluster',
                               group = 'exp')) + 
          geom_point(aes(shape = exp)) + scale_color_manual(values = color_pal2) +
          theme_bw() + ggtitle(paste0(temp.x, ' vs ', temp.y)))
}


```

**In general, we points clustering by cell type rather than by experiment, which is good. If we saw things clustering by experiment that would lead us to think there were batch effects.**

### Enrichment Effect

Using lines to see if there is an enrichment shift

```{r}

pca$group <- paste0(ifelse(pca$exp %in% c('enrMigr1','Migr1'), 'Migr1','Mpl'),pca$cluster)

for (i in 1:dim(combos)[2]){
  # print(combos[1,i])
  temp.x <- combos[1,i]
  # print(combos[2,i])
  temp.y <- combos[2,i]
  
  print(ggplot(pca, aes_string(x = temp.x, y = temp.y, color = 'cluster',
                               group = 'group')) + 
          geom_line(aes(group = group, color = cluster)) +
          geom_point(aes(shape = exp)) + scale_color_manual(values = color_pal2) +
          theme_bw() + ggtitle(paste0(temp.x, ' vs ', temp.y)))
}


ggplot(pca, aes_string(x = 'PC1', y = 'PC2', color = 'cluster',
                               group = 'group')) + 
  geom_line(aes(group = group, color = cluster)) +
  geom_point(aes(shape = exp)) + #scale_color_manual(values = color_pal2) +
  theme_bw() + ggtitle(paste0(temp.x, ' vs ', temp.y))
  
```

**In general, the lines are not all parrallel to each other. If they were, and headed in the same direciton (which isn't indicated on these lines), it would indicate that the enrichment process had a parallel effect on all samples, which could be a batch effect**

### Heatmap

```{r}
av.with.pca <- cbind(av, pc.vectors[,1:20])
temp.cor <- cor(av, method = 'spearman')
temp.cor.pca <- cor(av.with.pca, method = 'spearman')
```

#### Without PC Vectors

##### By Experiment

```{r}
col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-1' = 'purple', 'MK-2' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred'))



order.rows <- data.frame(coln = colnames(temp.cor))

order.rows$exp <- tstrsplit(order.rows$coln, '_', keep = 1)[[1]]
order.rows$ct <- tstrsplit(order.rows$coln, '_', keep = 2)[[1]]

order.rows <- order.rows[order(order.rows$exp, order.rows$ct),]$coln

# order.rows

temp.cor2 <- temp.cor[order.rows, order.rows]

key_ <- data.frame(ident.combo = colnames(temp.cor2))

rownames(key_) <- 1:length(colnames(av))

key_$exp <- tstrsplit(key_$ident.combo, '_', keep = 1)[[1]]
key_$cluster <- tstrsplit(key_$ident.combo, '_', keep = 2)[[1]]

column_ha = HeatmapAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)

row_ha = rowAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = key_$exp,
        column_split = key_$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

##### By Cluster

```{r}
order.rows <- data.frame(coln = colnames(temp.cor))

order.rows$exp <- tstrsplit(order.rows$coln, '_', keep = 2)[[1]]

order.rows <- order.rows[order(order.rows$exp),]$coln

temp.cor2 <- temp.cor2[order.rows, order.rows]

rownames(key_) <- key_$ident.combo
key_ <- key_[rownames(temp.cor2),]

column_ha = HeatmapAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)

row_ha = rowAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)

Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = key_$cluster,
        column_split = key_$cluster,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

#### With PC Vectors

##### By Experiment

```{r}
col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-1' = 'purple', 'MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1' = 'blue', 'Mpl' = 'red', 'enrMigr1' = 'violet',
                   'enrMpl' = 'darkred','PC' = 'black'))



order.rows <- data.frame(coln = colnames(temp.cor.pca))
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])

order.rows <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

order.rows

temp.cor2 <- temp.cor.pca[order.rows, order.rows]

key_ <- data.frame(ident.combo = colnames(temp.cor2))

rownames(key_) <- 1:length(colnames(av.with.pca))

key_$pc <- ifelse(key_$ident.combo %in% paste0('PC',1:20), 1,0)
key_$exp <- ifelse(key_$pc == T, 'PC',tstrsplit(key_$ident.combo, '_', keep = 1)[[1]])
key_$cluster <- ifelse(key_$pc == T, 'PC', tstrsplit(key_$ident.combo, '_', keep = 2)[[1]])

column_ha = HeatmapAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)

row_ha = rowAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = key_$exp,
        column_split = key_$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

##### By Celltype

```{r}
order.rows <- data.frame(coln = colnames(temp.cor.pca))

order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '_', keep = 2)[[1]])

# order.rows$exp <- tstrsplit(order.rows$coln, '_', keep = 2)[[1]]

order.rows <- order.rows[order(order.rows$pc, order.rows$exp, order.rows$ct),]$coln

temp.cor2 <- temp.cor.pca[order.rows, order.rows]

rownames(key_) <- key_$ident.combo
key_
key_ <- key_[rownames(temp.cor2),]

column_ha = HeatmapAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)

row_ha = rowAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)

Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = key_$cluster,
        column_split = key_$cluster,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```


## 3b Cluster Centroid Deltas

```{r}
av <- AverageExpression(wbm)$RNA
# head(av)

av <- as.data.frame(t(av))

av$exp <- tstrsplit(rownames(av), '_', keep = 1)[[1]]

av$ct <- tstrsplit(rownames(av), '_', keep = 2)[[1]]

outs <- split(av, f = av$ct)
genes <- colnames(outs[[1]])[!colnames(outs[[1]]) %in% c('ct','exp')]
out.df <- data.frame(na.row = rep(NA,length(genes)),
                     row.names = genes)

for (p in outs){
  temp.df <- t(p[,!(colnames(p) %in% c('ct','exp'))])
  ct <- tstrsplit(colnames(temp.df)[1], '_', keep = 2)[[1]]
  colnames(temp.df) <- tstrsplit(colnames(temp.df), '_', keep = 1)[[1]]
  
  combos <- combn(1:length(colnames(temp.df)), 2)
  
  if (dim(temp.df)[2] == 4){
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1','enrMpl')]
  }
  else{
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1')]
  }
  
  for (i in 1:dim(combos)[2]){
    clm1 <- combos[1,i]
    clm2 <- combos[2,i]
    # print(paste0(clm1, clm2))
    
    if (i == 1){
      new.df.temp <- data.frame(V1 = temp.df[,clm1] - temp.df[,clm2])
      
    }
    else{
      new.df.temp <- cbind(new.df.temp, temp.df[,clm1] - temp.df[,clm2])
    }
    colnames(new.df.temp)[i] <- paste0(colnames(temp.df)[clm1],
                                       '_vs_', colnames(temp.df)[clm2],
                                       '.',ct)
  }
  
  out.df <- cbind(out.df, new.df.temp)
  
}

# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 2)[[1]]))
# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 1)[[1]]))

out.df <- out.df[,-1]
```

### PCs

```{r}
pca <- prcomp(out.df)

pc.vectors <- pca$x

out.df.pca <- cbind(out.df, pc.vectors[,1:20])

temp.cor.pca <- cor(out.df.pca, method = 'spearman')

temp.cor.pca <- temp.cor.pca[,!(grepl('Prog', colnames(temp.cor.pca)))]
# unique(tstrsplit(rownames(temp.cor.pca),'\\.', keep = 1)[[1]])[1:12]
col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-1' = 'purple', 'MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue',
                   'Migr1_vs_enrMigr1' = 'red',
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred',
                   'Mpl_vs_enrMpl' = 'green',
                   'enrMigr1_vs_enrMpl' = "orange",
                   'PC' = 'black'))



order.rows <- data.frame(coln = colnames(temp.cor.pca))
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

order.rows

temp.cor2 <- temp.cor.pca[order.rows, order.rows]

key_ <- data.frame(ident.combo = colnames(temp.cor2))

rownames(key_) <- 1:(length(colnames(out.df.pca))-3)

key_$pc <- ifelse(key_$ident.combo %in% paste0('PC',1:20), 1,0)
key_$exp <- ifelse(key_$pc == T, 'PC',tstrsplit(key_$ident.combo, '\\.', keep = 1)[[1]])
key_$cluster <- ifelse(key_$pc == T, 'PC', tstrsplit(key_$ident.combo, '\\.', keep = 2)[[1]])

column_ha = HeatmapAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)

row_ha = rowAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = key_$exp,
        column_split = key_$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')

# key_
```

## 3c Cluster Centroid pairwise deltas within cluster with PCs of the 44 cluster centroids

The same data I show in 3b, but instead of using the PCs from the deltas, I'm going to use the PCs from the 44 centroids.

```{r}
av <- AverageExpression(wbm)$RNA
# head(av)

av <- as.data.frame(t(av))

av$exp <- tstrsplit(rownames(av), '_', keep = 1)[[1]]

av$ct <- tstrsplit(rownames(av), '_', keep = 2)[[1]]

outs <- split(av, f = av$ct)
genes <- colnames(outs[[1]])[!colnames(outs[[1]]) %in% c('ct','exp')]
out.df <- data.frame(na.row = rep(NA,length(genes)),
                     row.names = genes)

for (p in outs){
  temp.df <- t(p[,!(colnames(p) %in% c('ct','exp'))])
  ct <- tstrsplit(colnames(temp.df)[1], '_', keep = 2)[[1]]
  colnames(temp.df) <- tstrsplit(colnames(temp.df), '_', keep = 1)[[1]]
  
  combos <- combn(1:length(colnames(temp.df)), 2)
  
  if (dim(temp.df)[2] == 4){
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1','enrMpl')]
  }
  else{
    temp.df <- temp.df[,c('Migr1','Mpl','enrMigr1')]
  }
  
  for (i in 1:dim(combos)[2]){
    clm1 <- combos[1,i]
    clm2 <- combos[2,i]
    # print(paste0(clm1, clm2))
    
    if (i == 1){
      new.df.temp <- data.frame(V1 = temp.df[,clm1] - temp.df[,clm2])
      
    }
    else{
      new.df.temp <- cbind(new.df.temp, temp.df[,clm1] - temp.df[,clm2])
    }
    colnames(new.df.temp)[i] <- paste0(colnames(temp.df)[clm1],
                                       '_vs_', colnames(temp.df)[clm2],
                                       '.',ct)
  }
  
  out.df <- cbind(out.df, new.df.temp)
  
}

# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 2)[[1]]))
# summary(as.factor(tstrsplit(colnames(out.df), '\\.', keep = 1)[[1]]))

out.df <- out.df[,-1]
```

### All Genes

```{r}
# head(out.df)
av <- AverageExpression(wbm)$RNA
av.pc <- prcomp(av)
av.pc <- av.pc$x
av.pc <- av.pc[,1:20]
# dim(out.df)
# dim(av.pc)
```


```{r}

out.df.pca <- cbind(out.df, av.pc)

temp.cor.pca <- cor(out.df.pca, method = 'spearman')

temp.cor.pca <- temp.cor.pca[,!(grepl('Prog', colnames(temp.cor.pca)))]
# unique(tstrsplit(rownames(temp.cor.pca),'\\.', keep = 1)[[1]])[1:12]
col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-1' = 'purple', 'MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue',
                   'Migr1_vs_enrMigr1' = 'red',
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred',
                   'Mpl_vs_enrMpl' = 'green',
                   'enrMigr1_vs_enrMpl' = "orange",
                   'PC' = 'black'))



order.rows <- data.frame(coln = colnames(temp.cor.pca))
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

order.rows

temp.cor2 <- temp.cor.pca[order.rows, order.rows]

key_ <- data.frame(ident.combo = colnames(temp.cor2))

rownames(key_) <- 1:(length(colnames(out.df.pca))-3)

key_$pc <- ifelse(key_$ident.combo %in% paste0('PC',1:20), 1,0)
key_$exp <- ifelse(key_$pc == T, 'PC',tstrsplit(key_$ident.combo, '\\.', keep = 1)[[1]])
key_$cluster <- ifelse(key_$pc == T, 'PC', tstrsplit(key_$ident.combo, '\\.', keep = 2)[[1]])

column_ha = HeatmapAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col, show_legend = F)

row_ha = rowAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = key_$exp,
        column_split = key_$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation', 
        show_heatmap_legend = FALSE)

# key_
```

### HV Genes

Using the top 2K highly variable genes from our dataset

```{r}
# head(out.df)
hv.genes <- VariableFeatures(wbm)
# length(hv.genes)
av <- AverageExpression(wbm)$RNA
av <- av[hv.genes,]
av.pc <- prcomp(av)
av.pc <- av.pc$x
av.pc <- av.pc[,1:20]

```


```{r}

out.df.pca <- cbind(out.df[hv.genes,], av.pc)

temp.cor.pca <- cor(out.df.pca, method = 'spearman')

temp.cor.pca <- temp.cor.pca[,!(grepl('Prog', colnames(temp.cor.pca)))]
# unique(tstrsplit(rownames(temp.cor.pca),'\\.', keep = 1)[[1]])[1:12]
col = list(CellType = c('B-cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono/Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00", 
                        'T-cell' = 'red', 'B-cell Prog.' = '#CC79A7',
                        'MkP' = 'violet','MK-1' = 'purple', 'MK-2' = 'darkgreen',
                        'PC' = 'black'),
           Exp = c('Migr1_vs_Mpl' = 'blue',
                   'Migr1_vs_enrMigr1' = 'red',
                   'Migr1_vs_enrMpl' = 'violet',
                   'Mpl_vs_enrMigr1' = 'darkred',
                   'Mpl_vs_enrMpl' = 'green',
                   'enrMigr1_vs_enrMpl' = "orange",
                   'PC' = 'black'))



order.rows <- data.frame(coln = colnames(temp.cor.pca))
order.rows$pc <- ifelse(order.rows$coln %in% paste0('PC',1:20),
                        1,0)
order.rows$exp <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 1)[[1]])
order.rows$ct <- ifelse(order.rows$pc == T, 'PC', 
                         tstrsplit(order.rows$coln, '\\.', keep = 2)[[1]])

order.rows <- order.rows[order(order.rows$pc,order.rows$exp, order.rows$ct),]$coln

# order.rows

temp.cor2 <- temp.cor.pca[order.rows, order.rows]

key_ <- data.frame(ident.combo = colnames(temp.cor2))

rownames(key_) <- 1:(length(colnames(out.df.pca))-3)

key_$pc <- ifelse(key_$ident.combo %in% paste0('PC',1:20), 1,0)
key_$exp <- ifelse(key_$pc == T, 'PC',tstrsplit(key_$ident.combo, '\\.', keep = 1)[[1]])
key_$cluster <- ifelse(key_$pc == T, 'PC', tstrsplit(key_$ident.combo, '\\.', keep = 2)[[1]])

column_ha = HeatmapAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col, show_legend = F)

row_ha = rowAnnotation(CellType = key_$cluster,Exp = key_$exp,
                              col = col, show_legend = F)


Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = key_$exp,
        column_split = key_$exp,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation', 
        show_heatmap_legend = FALSE)

# key_
```

# 4 Cluster T-Score Correlation 

40xGenes Delta of pairwise cluster centroid

Getting the t-scores for all genes doing differential expression within a cluster comparing MigR1 to MPL. Then putting those into a table, so each column is a cluster and the rows are t-scores for genes. Then doing a correlation analysis to see if certain cell types have the same pattern.

```{r}
# de.genes <- data.frame(p_val = numeric(),
#                        avg_logFC = numeric(),
#                        pct.1 = numeric(),
#                        pct.2 = numeric(),
#                        p_val_adj = numeric(),
#                        cluster = character(),
#                        gene = character())
# 
# for (i in unique(wbm$figure.labels3)){
#   print(i)
#   temp.subset <- subset(wbm, figure.labels3 == i)
#   x <- FindMarkers(temp.subset,
#                    ident.1 = 'MigR1',
#                    ident.2 = 'MPL',
#                    logfc.threshold = 0,
#                    min.pct = 0,
#                    group.by = 'Experiment',
#                    test.use = 'MAST')
#   x$cluster <- i
#   x$gene <- rownames(x)
#   x <- x[order(x$avg_log2FC, decreasing = T),]
#   de.genes <- rbind(de.genes,x)
# }
# 
# summary(as.factor(de.genes$cluster))
# write.table(de.genes,'./data/v4/mostly.all.genes.deg.clusters.txt',
#             sep = '\t')

x <- read.table('./data/v4/mostly.all.genes.deg.clusters.txt', sep = '\t')
dim(x)

# Limiting to only genes that are across all clusters

# Filter(function (elem) length(which(x$gene == elem)) == 11, x$gene)

freq <- table(x$gene)

genes.across.all <- names(freq[freq == 11])
class(genes.across.all)

length(genes.across.all)


x2 <- x[x$gene %in% genes.across.all,]

summary(as.factor(x2$cluster))

x2 <- x2[,c(1,6:7)]
x2.wide <- spread(x2, cluster, p_val)


rownames(x2.wide) <- x2.wide$gene

x2.wide <- x2.wide[,-1]

# rowSums(x2.wide) == 11

x3 <- x2.wide[rowSums(x2.wide) != 11,]

colSums(x3)

temp.cor <- cor(x3)
temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value, label = round(value,2))) +
  geom_tile() + geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('P-value Correlation') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .15) +
  theme(axis.text.x = element_text(angle = 90))
  
```

## 4a T-score Correlation 11x11

Doing combined enriched and non-enriched

### All Genes

```{r}
# wbm@assays$RNA@scale.data[1:10,1:10]

# cnts <- wbm@assays$RNA@scale.data
# t.scores <- as.data.frame(matrix(ncol = 11, nrow = dim(cnts)[1]))
# rownames(t.scores) <- rownames(cnts)
# cnt <- 1
# 
# for (cluster in unique(wbm$figure.labels3)){
#   print(cluster)
#   
#   colnames(t.scores)[cnt] <- cluster
#   cnt <- cnt + 1
#   
#   migr1.cells <- colnames(subset(wbm, figure.labels3 == cluster & Experiment == 'MigR1'))
#   mpl.cells <- colnames(subset(wbm, figure.labels3 == cluster & Experiment == 'MPL'))
#   
#   migr1.cnts <- cnts[,colnames(cnts) %in% migr1.cells]
#   mpl.cnts <- cnts[,colnames(cnts) %in% mpl.cells]
#   
#   t.score <- vector()
#   
#   for (i in 1:dim(migr1.cnts)[1]){
#     # print(i)
#     if (min(mpl.cnts[i,]) != max(mpl.cnts[i,]) | min(migr1.cnts[i,]) != max(migr1.cnts[i,])){
#       p <- t.test(migr1.cnts[i,], mpl.cnts[i,])
#       
#       t.score[i] <- p$statistic
#     }
#     else{
#       t.score[i] <- NA
#     }
#     if (i %% 1000 == 0){
#       print(i)
#     }
#   }
#   t.scores[,cluster] <- t.score
# }
# 
# 
# write.table(t.scores, './data/v4/t.scores.11.by.11.txt', sep = '\t', quote = F,
#             row.names = T)

t.scores <- read.table('./data/v4/t.scores.11.by.11.txt', sep = '\t')


complete.t.scores <- complete.cases(t.scores)

sum(complete.t.scores)

t.scores2 <- t.scores[complete.t.scores,]
?cor
spearman.cor <- cor(t.scores2, method = 'spearman')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('T-statistics Correlation (Spearman) All Genes') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
  theme(axis.text.x = element_text(angle = 90))
```

#### Using Pairwise-complete observations

Instead of removing any row that has NA values, we are going to use the the "use = pairwise.complete.obs" tag in the correlation.

```{r}
spearman.cor <- cor(t.scores, method = 'spearman', use = 'pairwise.complete.obs')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(round(temp.cor2$value,2) == 1, 0, temp.cor2$value)

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('T-statistics Correlation (Spearman) All Genes') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
  theme(axis.text.x = element_text(angle = 90))
```



### Using HV Genes Only

```{r}
t.scores <- read.table('./data/v4/t.scores.11.by.11.txt', sep = '\t')

complete.t.scores <- complete.cases(t.scores)

sum(complete.t.scores)

t.scores2 <- t.scores[complete.t.scores,]

summary(rownames(t.scores2) %in% VariableFeatures(wbm))

t.scores2 <- t.scores2[rownames(t.scores2) %in% VariableFeatures(wbm),]

spearman.cor <- cor(t.scores2, method = 'spearman')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('T-statistics Correlation (Spearman) HV Genes') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
  theme(axis.text.x = element_text(angle = 90))
```

#### Using Pairwise-complete observations

Instead of removing any row that has NA values, we are going to use the the "use = pairwise.complete.obs" tag in the correlation.

```{r}
t.scores <- t.scores[rownames(t.scores) %in% VariableFeatures(wbm),]
dim(t.scores)

spearman.cor <- cor(t.scores, method = 'spearman', use = 'pairwise.complete.obs')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(round(temp.cor2$value,2) == 1, 0, temp.cor2$value)

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('T-statistics Correlation (Spearman) All Genes') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
table(wbm$Condition3, wbm$figure.labels3)
```

## 4b Pairwise Comparison of T-Scores

```{r}
# wbm@assays$RNA@scale.data[1:10,1:10]
# 
# cnts <- wbm@assays$RNA@scale.data
# t.scores <- as.data.frame(matrix(ncol = 56, nrow = dim(cnts)[1]))
# rownames(t.scores) <- rownames(cnts)
# cnt <- 1
# 
# wbm$Condition.short <- factor(wbm$Condition, labels = c('eWT','eMUT','WT','MUT'))
# 
# combos <- combn(1:4,2)
# 
# clusters <- unique(wbm$figure.labels3)
# clusters <- clusters[-c(2,8)]
# 
# for (cluster in clusters){
#   print(cluster)
#   
#   WT.cells <- colnames(subset(wbm, 
#                               figure.labels3 == cluster & Condition.short == 'WT'))
#   MUT.cells <- colnames(subset(wbm, 
#                                figure.labels3 == cluster & Condition.short == 'MUT'))
#   eWT.cells <- colnames(subset(wbm, 
#                                figure.labels3 == cluster & Condition.short == 'eWT'))
#   eMUT.cells <- colnames(subset(wbm, 
#                                 figure.labels3 == cluster & Condition.short == 'eMUT'))
#   
#   WT.counts <- cnts[,colnames(cnts) %in% WT.cells]
#   MUT.counts <- cnts[,colnames(cnts) %in% MUT.cells]
#   eWT.counts <- cnts[,colnames(cnts) %in% eWT.cells]
#   eMUT.counts <- cnts[,colnames(cnts) %in% eMUT.cells]
#   
#   temp <- list(WT.counts, MUT.counts, eWT.counts, eMUT.counts)
#   names(temp) <- c('WT.counts', 'MUT.counts', 'eWT.counts', 'eMUT.counts')
#   
#   for (comb in 1:dim(combos)[2]){
#     cnt1 <- combos[1,comb]
#     cnt2 <- combos[2,comb]
#     name1 <- tstrsplit(names(temp)[cnt1], '\\.', keep = 1)[[1]]
#     name2 <- tstrsplit(names(temp)[cnt2], '\\.', keep = 1)[[1]]
#     
#     cnts1 <- temp[[cnt1]]
#     cnts2 <- temp[[cnt2]]
#     
#     
#     clm.name <- paste0(cluster,'_', name1,'.vs.',name2)
#     print(clm.name)
#     
#     t.score <- vector()
#   
#     for (i in 1:dim(cnts1)[1]){
#       
#       # print(i)
#       if (min(cnts1[i,]) != max(cnts1[i,]) | min(cnts2[i,]) != max(cnts2[i,])){
#         p <- t.test(cnts1[i,], cnts2[i,])
#         
#         t.score[i] <- p$statistic
#       }
#       else{
#         t.score[i] <- NA
#       }
#       if (i %% 1000 == 0){
#         print(i)
#       }
#     }
#     t.scores[,clm.name] <- t.score
#   }
# }
# 
# t.scores
# 
# t.scores.save <- t.scores
# 
# t.scores <- t.scores[,57:110]
# 
# t.scores
# 
# write.table(t.scores,'./data/v4/t.scores.54.by.54.txt', quote = F, sep = '\t', row.names = T)

t.scores <- read.table('./data/v4/t.scores.54.by.54.txt', sep = '\t')

```

Looking at the missingness

```{r}

nas <- data.frame(ncol = 1)
colnames(nas) <- 'Total NAs'

for (i in 1:54){
  nas[i,1] <- sum(is.na(t.scores[,i]))
  rownames(nas)[i] <- colnames(t.scores)[i]
}
nas$ct <- tstrsplit(rownames(nas),'_', keep = 1)[[1]]
nas$comp <- tstrsplit(rownames(nas),'_', keep = 2)[[1]]

nas$pct.missing <- nas$`Total NAs` / dim(t.scores)[1]

# dim(t.scores)

ggplot(nas, aes(x = ct, y = comp, label = `Total NAs`, fill = pct.missing)) +
        geom_tile() + geom_text() + 
        scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', 
                             midpoint = .5) + ggtitle('Missingness') +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### PC Plot of T-Scores

```{r}
t.scores2 <- t.scores[complete.cases(t.scores),]
pca <- prcomp(t.scores2)
t.pca.vectors <- pca$x
pca.sd <- pca$sdev
pca <- as.data.frame(pca$rotation)

pca.sd <- data.frame(pc = 1:54,
                     sd = pca.sd)

ggplot(pca.sd[1:20,], aes(x = pc, y = sd)) + geom_point() + geom_line() + theme_bw()
# colnames(pca) <- paste0(colnames(pca),' (',round(pca.sd,2),')')

pca$cluster <- tstrsplit(rownames(pca),'_', keep = 1)[[1]]
pca$comparison <- tstrsplit(rownames(pca),'_', keep = 2)[[1]]
pca$Grp1 <- tstrsplit(pca$comparison, '\\.', keep = 1)[[1]]
pca$Grp2 <- tstrsplit(pca$comparison, '\\.', keep = 3)[[1]]
# pca

combos <- combn(paste0('PC',1:5),2)
# combos

# colnames(pca)[1:5] <- paste0(colnames(pca)[1:5], ' (SD = ', pca.sd[1:5], ')')

for (i in 1:dim(combos)[2]){
  # print(combos[1,i])
  temp.x <- combos[1,i]
  # print(combos[2,i])
  temp.y <- combos[2,i]
  
  print(ggplot(pca, aes_string(x = temp.x, y = temp.y, color = 'cluster',
                               group = 'comparison')) + 
                geom_point(aes(shape = comparison)) + 
                scale_color_manual(values = color_pal2) +
                theme_bw() + ggtitle(paste0(temp.x, ' vs ', temp.y)))
}
```


### Heatmap w/o PC

##### Ordered by Cluster (Complete Obs)

Only looking at complete observations

```{r}
complete.t.scores <- complete.cases(t.scores)

sum(complete.t.scores)

t.scores2 <- t.scores[complete.t.scores,]

spearman.cor <- cor(t.scores2, method = 'spearman')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + # geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('T-statistics Correlation (Spearman)') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

```{r}
library(ComplexHeatmap)
# 
# set.seed(123)
# nr1 = 4; nr2 = 8; nr3 = 6; nr = nr1 + nr2 + nr3
# nc1 = 6; nc2 = 8; nc3 = 10; nc = nc1 + nc2 + nc3
# mat = cbind(rbind(matrix(rnorm(nr1*nc1, mean = 1,   sd = 0.5), nr = nr1),
#           matrix(rnorm(nr2*nc1, mean = 0,   sd = 0.5), nr = nr2),
#           matrix(rnorm(nr3*nc1, mean = 0,   sd = 0.5), nr = nr3)),
#     rbind(matrix(rnorm(nr1*nc2, mean = 0,   sd = 0.5), nr = nr1),
#           matrix(rnorm(nr2*nc2, mean = 1,   sd = 0.5), nr = nr2),
#           matrix(rnorm(nr3*nc2, mean = 0,   sd = 0.5), nr = nr3)),
#     rbind(matrix(rnorm(nr1*nc3, mean = 0.5, sd = 0.5), nr = nr1),
#           matrix(rnorm(nr2*nc3, mean = 0.5, sd = 0.5), nr = nr2),
#           matrix(rnorm(nr3*nc3, mean = 1,   sd = 0.5), nr = nr3))
#    )
# mat = mat[sample(nr, nr), sample(nc, nc)] # random shuffle rows and columns
# rownames(mat) = paste0("row", seq_len(nr))
# colnames(mat) = paste0("column", seq_len(nc))
# 
# Heatmap(mat)
# 
# x.temp <- as.matrix(temp.cor2[])
# Heatmap(temp.cor)
```


```{r}
row.anno <- data.frame(rownames = row.names(temp.cor))

head(row.anno)
row.anno$ct <- tstrsplit(row.anno$rownames,'_', keep = 1)[[1]]
row.anno$type1 <- tstrsplit(tstrsplit(row.anno$rownames,'_', keep = 2)[[1]],'\\.', 
                         keep = 1)[[1]] 
row.anno$type2 <- tstrsplit(tstrsplit(row.anno$rownames,'_', keep = 2)[[1]],'\\.', 
                         keep = 3)[[1]] 

col = list(CellType = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'black'),
           Grp1 = c('WT' = 'blue','eWT' = 'darkblue','MUT' = 'red'),
           Grp2 = c('eWT' = 'darkblue','MUT' = 'red', 'eMUT' = 'darkred'))

column_ha = HeatmapAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)

row_ha = rowAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)

# temp.cor[temp.cor == 1] <- 0
Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = F, cluster_columns = F,
        split = row.anno$ct,
        column_split = row.anno$ct,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')
```

#### Ordered by Groupings

```{r}
temp.df <- data.frame(names = colnames(temp.cor))

temp.df$group <- tstrsplit(temp.df$names, '_', keep = 2)[[1]]

temp.df <- temp.df[order(temp.df$group),]

order_ <- as.numeric(rownames(temp.df))

# order_

temp.cor2 <- temp.cor[order_, order_]
# temp.cor2


row.anno <- data.frame(rownames = row.names(temp.cor2))

head(row.anno)
row.anno$ct <- tstrsplit(row.anno$rownames,'_', keep = 1)[[1]]
row.anno$type1 <- tstrsplit(tstrsplit(row.anno$rownames,'_', keep = 2)[[1]],'\\.', 
                         keep = 1)[[1]] 
row.anno$type2 <- tstrsplit(tstrsplit(row.anno$rownames,'_', keep = 2)[[1]],'\\.', 
                         keep = 3)[[1]] 

row.anno$combo <- tstrsplit(row.anno$rownames, '_', keep = 2)[[1]]

col = list(CellType = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'black'),
           Grp1 = c('WT' = 'blue','eWT' = 'darkblue','MUT' = 'red'),
           Grp2 = c('eWT' = 'darkblue','MUT' = 'red', 'eMUT' = 'darkred'))

column_ha = HeatmapAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)

row_ha = rowAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)
Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = F, cluster_columns = F,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')

col = list(CellType = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'black'),
           Comparison = c('WT.vs.MUT' = 'red', 'WT.vs.eMUT' = 'orange',
                          'WT.vs.eWT' = 'yellow', 'MUT.vs.eMUT' = 'green',
                          'MUT.vs.eWT' = 'blue', 'eWT.vs.eMUT' = 'violet'))

column_ha = HeatmapAnnotation(CellType = row.anno$ct,Comparison = row.anno$combo,
                              col = col)

row_ha = rowAnnotation(CellType = row.anno$ct, Comparison = row.anno$combo,
                       col = col)

Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = row.anno$combo,
        column_split = row.anno$combo,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')


```


### Heatmap w/ PC

##### Ordered by Cluster (Complete Obs)

Only looking at complete observations

```{r}
complete.t.scores <- complete.cases(t.scores)

sum(complete.t.scores)

t.scores2 <- t.scores[complete.t.scores,]
t.scores2 <- cbind(t.scores2, t.pca.vectors[rownames(t.scores2),1:20])

spearman.cor <- cor(t.scores2, method = 'spearman')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + # geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('T-statistics Correlation (Spearman)') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())
```


```{r}
row.anno <- data.frame(rownames = row.names(temp.cor))

# head(row.anno)
row.anno$pc <- ifelse(row.anno$rownames %in% paste0('PC',1:20), 1,0)
row.anno$ct <- ifelse(row.anno$pc == T, 'PC', 
                      tstrsplit(row.anno$rownames,'_', keep = 1)[[1]])
row.anno$type1 <- ifelse(row.anno$pc == T, 'PC', 
                         tstrsplit(tstrsplit(row.anno$rownames,'_', 
                                             keep = 2)[[1]],'\\.', 
                         keep = 1)[[1]])
row.anno$type2 <- ifelse(row.anno$pc == T, 'PC', 
                         tstrsplit(tstrsplit(row.anno$rownames,'_', 
                                             keep = 2)[[1]],'\\.', 
                         keep = 3)[[1]])

col = list(CellType = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'PC' = 'black'),
           Grp1 = c('WT' = 'blue','eWT' = 'darkblue','MUT' = 'red', 'PC' = 'black'),
           Grp2 = c('eWT' = 'darkblue','MUT' = 'red', 'eMUT' = 'darkred', 
                    'PC' = 'black'))

column_ha = HeatmapAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)

row_ha = rowAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)

# temp.cor[temp.cor == 1] <- 0
Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = F, cluster_columns = F,
        split = row.anno$ct,
        column_split = row.anno$ct,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')
```

#### Ordered by Groupings

```{r}
temp.df <- data.frame(names = colnames(temp.cor))

temp.df$group <- ifelse(temp.df$names %in% paste0('PC', 1:20), 'PC',
                        tstrsplit(temp.df$names, '_', keep = 2)[[1]])

temp.df <- temp.df[order(temp.df$group),]
temp.df <- rbind(temp.df[temp.df$group != 'PC',], temp.df[temp.df$group == 'PC',])

order_ <- as.numeric(rownames(temp.df))

# order_

temp.cor2 <- temp.cor[order_, order_]
# temp.cor2


row.anno <- data.frame(rownames = row.names(temp.cor2))

head(row.anno)
row.anno$pc <- ifelse(row.anno$rownames %in% paste0('PC',1:20), 1, 0)
row.anno$ct <- ifelse(row.anno$pc == 1, 'PC',
                      tstrsplit(row.anno$rownames,'_', keep = 1)[[1]])
row.anno$type1 <- ifelse(row.anno$pc == 1, 'PC',
                         tstrsplit(tstrsplit(row.anno$rownames,'_', 
                                             keep = 2)[[1]],'\\.', 
                         keep = 1)[[1]])
row.anno$type2 <- ifelse(row.anno$pc == 1, 'PC',
                         tstrsplit(tstrsplit(row.anno$rownames,'_', 
                                             keep = 2)[[1]],'\\.', 
                         keep = 3)[[1]])

row.anno$combo <- ifelse(row.anno$pc == 1, 'PC',
                         tstrsplit(row.anno$rownames, '_', keep = 2)[[1]])

col = list(CellType = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'PC' = 'black'),
           Grp1 = c('WT' = 'blue','eWT' = 'darkblue','MUT' = 'red', 'PC' = 'black'),
           Grp2 = c('eWT' = 'darkblue','MUT' = 'red', 'eMUT' = 'darkred', 
                    'PC' = 'black'))

column_ha = HeatmapAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)

row_ha = rowAnnotation(CellType = row.anno$ct, Grp1 = row.anno$type1,
                           Grp2 = row.anno$type2, col = col)
Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = F, cluster_columns = F,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')

col = list(CellType = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'PC' = 'black'),
           Comparison = c('WT.vs.MUT' = 'red', 'WT.vs.eMUT' = 'orange',
                          'WT.vs.eWT' = 'yellow', 'MUT.vs.eMUT' = 'green',
                          'MUT.vs.eWT' = 'blue', 'eWT.vs.eMUT' = 'violet',
                          'PC' = 'black'))

column_ha = HeatmapAnnotation(CellType = row.anno$ct,Comparison = row.anno$combo,
                              col = col)

row_ha = rowAnnotation(CellType = row.anno$ct, Comparison = row.anno$combo,
                       col = col)

Heatmap(temp.cor2, 
        top_annotation = column_ha, right_annotation = row_ha,
        split = row.anno$combo,
        column_split = row.anno$combo,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```


## 4c Cluster Intra T-Scores

```{r}
# cnts <- wbm@assays$RNA@scale.data
# dim(combn(unique(Idents(wbm)), 2))
# combos <- combn(unique(Idents(wbm)),2)
# t.scores <- as.data.frame(matrix(ncol = dim(combos)[2], nrow = dim(cnts)[1]))
# rownames(t.scores) <- rownames(cnts)
# cnt <- 1
# 
# for (i in 1:dim(combos)[2]){
#   ct1 <- combos[1,i]
#   ct2 <- combos[2,i]
# 
#   comparison <- paste0(ct1, '_vs_', ct2)
# 
#   print(paste0(i,'/',dim(combos)[2],' ',comparison))
# 
#   colnames(t.scores)[cnt] <- comparison
#   cnt <- cnt + 1
# 
#   ct1.cells <- colnames(subset(wbm, figure.labels3 == ct1))
#   ct2.cells <- colnames(subset(wbm, figure.labels3 == ct2))
# 
#   ct1.cnts <- cnts[,colnames(cnts) %in% ct1.cells]
#   ct2.cnts <- cnts[,colnames(cnts) %in% ct2.cells]
# 
#   t.score <- vector()
# 
#   for (i in 1:dim(ct1.cnts)[1]){
#     # print(i)
#     if (min(ct1.cnts[i,]) != max(ct1.cnts[i,]) |
#         min(ct2.cnts[i,]) != max(ct2.cnts[i,])){
#       p <- t.test(ct1.cnts[i,], ct2.cnts[i,])
# 
#       t.score[i] <- p$statistic
#     }
#     else{
#       t.score[i] <- NA
#     }
#     if (i %% 5000 == 0){
#       print(i)
#     }
#   }
#   t.scores[,comparison] <- t.score
# }
# 
# 
# write.table(t.scores, './data/v4/t.scores.all.between.cluster.txt',
#             sep = '\t', quote = F,
#             row.names = T)
```



```{r}
between.clusters <- read.table('./data/v4/t.scores.all.between.cluster.txt',
                               sep = '\t')

between.clusters

t.scores <- read.table('./data/v4/t.scores.11.by.11.txt', sep = '\t')

t.scores2 <- cbind(t.scores, between.clusters)
```

### All Genes

```{r}
spearman.cor <- cor(t.scores2, method = 'spearman', 
                    use = "pairwise.complete.obs")

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)

# ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
#   geom_tile() + geom_text() + 
#   xlab(NULL) + ylab(NULL) + ggtitle ('T-statistics Correlation (Spearman) All Genes') +
#   scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
#   theme(axis.text.x = element_text(angle = 90))
```

```{r}
Heatmap(temp.cor)
```


```{r}
temp.df <- data.frame(names = colnames(temp.cor))
temp.df$group <- ifelse(grepl('_',temp.df$names), 'comparison', 'singular')
temp.df$ct1 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 1)[[1]])
temp.df$ct2 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 2)[[1]])

col = list(CellType1 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'),
           CellType2 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'))

column_ha = HeatmapAnnotation(CellType1 = temp.df$ct1, CellType2 = temp.df$ct2,
                              col = col)

row_ha = rowAnnotation(CellType1 = temp.df$ct1, CellType2 = temp.df$ct2,
                              col = col)
Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = T, cluster_columns = T,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')

Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        # split = row.anno$combo,
        # column_split = row.anno$combo,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

### Highly Variable Genes

```{r}
spearman.cor <- cor(t.scores2[hv.genes,], method = 'spearman', 
                    use = "pairwise.complete.obs")

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)
```

```{r}
temp.df <- data.frame(names = colnames(temp.cor))
temp.df$group <- ifelse(grepl('_',temp.df$names), 'comparison', 'singular')
temp.df$ct1 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 1)[[1]])
temp.df$ct2 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 2)[[1]])

col = list(CellType1 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'),
           CellType2 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'))

column_ha = HeatmapAnnotation(CellType1 = temp.df$ct1, CellType2 = temp.df$ct2,
                              col = col, show_legend = F)

row_ha = rowAnnotation(CellType1 = temp.df$ct1, CellType2 = temp.df$ct2,
                              col = col, show_legend = F)
Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = T, cluster_columns = T,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')

Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        # split = row.anno$combo,
        # column_split = row.anno$combo,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')
```

###Subsetting to MK Lineages

#### All Genes

```{r}
spearman.cor <- cor(t.scores2[,], method = 'spearman', 
                    use = "pairwise.complete.obs")

temp.cor <- as.data.frame(spearman.cor)

mk.lineages <- c('MEP','MkP','MK.1','MK.2')
temp.cor <- temp.cor[mk.lineages,]
```

```{r}
temp.df <- data.frame(names = colnames(temp.cor))
temp.df$group <- ifelse(grepl('_',temp.df$names), 'comparison', 'singular')
temp.df$ct1 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 1)[[1]])
temp.df$ct2 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 2)[[1]])

temp.df2 <- data.frame(names = rownames(temp.cor))
temp.df2$group <- ifelse(grepl('_',temp.df2$names), 'comparison', 'singular')
temp.df2$ct1 <- ifelse(temp.df2$group == 'singular', temp.df2$names,
                      tstrsplit(temp.df2$names,'_vs_', keep = 1)[[1]])
temp.df2$ct2 <- ifelse(temp.df2$group == 'singular', temp.df2$names,
                      tstrsplit(temp.df2$names,'_vs_', keep = 2)[[1]])

col = list(CellType1 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'),
           CellType2 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'))

column_ha = HeatmapAnnotation(CellType1 = temp.df$ct1, CellType2 = temp.df$ct2,
                              col = col, show_legend = F)

row_ha = rowAnnotation(CellType1 = temp.df2$ct1, CellType2 = temp.df2$ct2,
                              col = col, show_legend = T)
Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = T, cluster_columns = T,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')

Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        # split = row.anno$combo,
        # column_split = row.anno$combo,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')

```

#### HV Genes Genes

```{r}
spearman.cor <- cor(t.scores2[hv.genes,], method = 'spearman', 
                    use = "pairwise.complete.obs")

temp.cor <- as.data.frame(spearman.cor)

mk.lineages <- c('MEP','MkP','MK.1','MK.2')
temp.cor <- temp.cor[mk.lineages,]
```

```{r}
temp.df <- data.frame(names = colnames(temp.cor))
temp.df$group <- ifelse(grepl('_',temp.df$names), 'comparison', 'singular')
temp.df$ct1 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 1)[[1]])
temp.df$ct2 <- ifelse(temp.df$group == 'singular', temp.df$names,
                      tstrsplit(temp.df$names,'_vs_', keep = 2)[[1]])

temp.df2 <- data.frame(names = rownames(temp.cor))
temp.df2$group <- ifelse(grepl('_',temp.df2$names), 'comparison', 'singular')
temp.df2$ct1 <- ifelse(temp.df2$group == 'singular', temp.df2$names,
                      tstrsplit(temp.df2$names,'_vs_', keep = 1)[[1]])
temp.df2$ct2 <- ifelse(temp.df2$group == 'singular', temp.df2$names,
                      tstrsplit(temp.df2$names,'_vs_', keep = 2)[[1]])

col = list(CellType1 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'),
           CellType2 = c('B.cell' = '#009E73', 'CMP' = '#CC79A7',
                        'Granulocyte' = '#0072B2', 'Mono.Macro' = '#E69F00',
                        'Erythroid' = '#56B4E9','MEP' = "#D55E00",
                        'MkP' = 'violet','MK.1' = 'purple', 'MK.2' = 'darkgreen',
                        'T.cell' = 'black', 'B.cell.Prog.' = 'orange'))

column_ha = HeatmapAnnotation(CellType1 = temp.df$ct1, CellType2 = temp.df$ct2,
                              col = col, show_legend = F)

row_ha = rowAnnotation(CellType1 = temp.df2$ct1, CellType2 = temp.df2$ct2,
                              col = col, show_legend = T)
Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        cluster_rows = T, cluster_columns = T,
        row_names_gp = gpar(fontsize = 0),
        column_names_gp = gpar(fontsize = 0),
        name = 'Correlation')

Heatmap(temp.cor, 
        top_annotation = column_ha, right_annotation = row_ha,
        # split = row.anno$combo,
        # column_split = row.anno$combo,
        show_column_names = F, show_row_names = F,
        cluster_rows = F, cluster_columns = F,
        name = 'Correlation')

```

Looking to see if MK-1 vs MK-2 has the same pattern as within MkP (MigR1 vs MPL) has the same overall correlation. Looking at the pink-pink (within MkP) row, we do not see a high correlation with with the Purple/Green Column, so this doesn't hold true here.

# 5 GO Term w/in clusters

Using (LRPath) [http://lrpath.ncibi.org] from Maureen Sartor's lab. Should this just been done with WT? Could we also do this with just MUT?

Need to do unbiased DE expression, not using any cutoffs. Looking within all clusters comparing WT vs MUT (normal WBM and enriched combined). This data will be inputted into the online interface, and then also used within the next section.

## DE

```{r}
# de.genes <- data.frame(p_val = numeric(),
#                        avg_logFC = numeric(),
#                        pct.1 = numeric(),
#                        pct.2 = numeric(),
#                        p_val_adj = numeric(),
#                        cluster = character(),
#                        gene = character())
# 
# wbm@meta.data
# 
# for (i in unique(wbm$figure.labels3)){
#   print(i)
#   temp.subset <- subset(wbm, figure.labels3 == i)
#   x <- FindMarkers(temp.subset,
#                    ident.1 = 'MigR1',
#                    ident.2 = 'MPL',
#                    logfc.threshold = 0,
#                    min.pct = 0,
#                    group.by = 'Experiment')
#   x$cluster <- i
#   x$gene <- rownames(x)
#   x <- x[order(x$avg_log2FC, decreasing = T),]
#   de.genes <- rbind(de.genes,x)
# }
# 
# summary(as.factor(de.genes$cluster))
# # table(rownames(de.genes))
# 



# write.table(de.genes, './data/v4/genes.for.LR.Path.within.cluster.all.txt', sep = '\t')
```

## Exporting Results to LRPath Format

```{r}
# x <- read.table('./data/v4/genes.for.LR.Path.within.cluster.all.txt', sep = '\t')
# 
# mus.ref <- data.table::fread('../mus_mus_gene_id.txt', header = T)
# colnames(mus.ref)[2] <- 'gene'
# 
# 
# out <- split(x, f = x$cluster)
# 
# for (i in 1:length(out)){
#   temp.name <- unique(out[[i]]$cluster)
#   temp.name <- gsub('/','-',temp.name)
#   
#   print(temp.name)
#   
#   temp <- out[[i]]
#   print(summary(temp$gene %in% mus.ref$gene))
#   
#   temp.df <- merge(temp, mus.ref, by = 'gene')
#   
#   temp.df$entrez.name <- paste0('mmu:', temp.df$GeneID)
#   
#   temp.df <- temp.df[,c('GeneID','p_val_adj','avg_log2FC')]
#   
#   write.table(temp.df, paste0('./data/v4/LRPath.gene.lists/',
#                               temp.name,'.txt'),
#               sep = '\t', col.names = F, quote = F, row.names = F)
# }
```


```{r}
# mkp.degs <- de.genes[de.genes$cluster == 'MkP',]
# mkp.degs
# mus.ref <- read.table('../mus_mus_gene_id.txt', sep = '\t',header = T)
# mus.ref2 <- data.table::fread('../mus_mus_gene_id.txt', header = T)
# 
# summary(mkp.degs$gene %in% mus.ref2$Symbol)
# 
# missing.genes <- mkp.degs$gene[!(mkp.degs$gene %in% mus.ref2$Symbol)]
# 
# missing.genes
# 
# mus.ref2$Symbol[grepl('sept', mus.ref2$Symbol, ignore.case = T)] # H2afv
# 
# colnames(mus.ref2)[2] <- 'gene'
# 
# df <- merge(mkp.degs, mus.ref2, by = 'gene')
# 
# df$entrez.name <- paste0('mmu:',df$GeneID)
# 
# df <- df[,c('GeneID','p_val_adj','avg_log2FC')]
# 
# write.table(df, './data/v4/LRPath.mkps.txt', sep = '\t', 
#             col.names = F, quote = F, row.names = F)
```


## Cleaning Files

```{r}
library(data.table)
bcell.prog <-fread('./data/v4/LRPath.results/within.BcellProg.txt', sep = '\t')
bcell.prog <- bcell.prog[bcell.prog$FDR < 0.05,]

files <- list.files('./data/v4/LRPath.results/')
files <- files[grepl('txt', files)]

cnt <- 1
for (i in files){
  # print(i)
  temp <- fread(paste0('./data/v4/LRPath.results/',i), sep = '\t')
  
  temp$cluster <- tstrsplit(i,'\\.', keep = 2)[[1]]
  if (cnt == 1){
    LR.results <- temp
    cnt <- 0
  }
  else{
    LR.results <- rbind(LR.results, temp)
  }
}

# summary(as.factor(LR.results$cluster))

LR.sig.results <- LR.results[LR.results$FDR < 0.05,] 

# summary(as.factor(LR.sig.results$cluster))

# table(LR.sig.results$cluster, LR.sig.results$ConceptType)

LR.sig.results$Direction2 <- ifelse(LR.sig.results$Direction == 'down', 'up','down')
```

## Clusters

Looking at clusters individually

### MkP

```{r}
mkp <- LR.sig.results[LR.sig.results$cluster == 'MkP',]

mkp[mkp$ConceptType == 'GOBP',]
mkp[mkp$ConceptType == 'KEGG',]
```

```{r}
splt.obj <- tstrsplit(mkp[mkp$ConceptType == 'GOBP',]$Name,' ')

words <- c()

for (i in 1:length(splt.obj)){
  words <- c(words,splt.obj[[i]])
}

words <- words[!is.na(words)]
# length(words)

# summary(as.factor(words))
```

## 5a Correlation of GO Terms

Reading in all the Terms

```{r}
# setwd('data/v4/LRPath.results/')
files = list.files('./data/v4/LRPath.results/')
files <- files[grepl('.txt',files)]
# files

cnt <- 1

for (i in files){
  # print(i)
  x <- fread(paste0('./data/v4/LRPath.results/',i))
  x$cluster <- tstrsplit(i,'\\.', keep = 2)[[1]]
  
  if (cnt == 1){
    go.df <- x
    cnt <- 2
  }
  else{
    go.df <- rbind(go.df, x)
  }
}

# summary(as.factor(go.df$cluster))
# summary(as.factor(go.df$ConceptType))
```

Subsetting to just GOBP

```{r}
go.df <- go.df[go.df$ConceptType == 'GOBP',]

tmp <- table(go.df$Name, go.df$cluster)

# summary(as.factor(go.df$Name))

go.bps <- rownames(tmp[rowSums(tmp) == 9,])

# length(go.bps)
```

### Odds Ratio

```{r}
out <- split(go.df, f = go.df$cluster)

# w.go.df

cnt = 1 

for (i in 1:length(out)){
  temp <- as.data.frame(out[[i]])
  rown <- as.character(temp$Name)
  coln <- paste0(names(out)[i],'_',colnames(temp)[5])
  colnames(temp)[5] <- coln
  temp <- data.frame(temp[,5])
  rownames(temp) <- rown
  temp <- data.frame(temp[go.bps,])
  colnames(temp) <- coln
  
  if(cnt == 1){
    w.go.df <- temp
    cnt = 2
  }
  else{
    w.go.df <- cbind(w.go.df, temp)
  }
  
}
```

```{r}
# summary(as.factor(go.df$cluster))
# length(go.bps)
```


```{r}
spearman.cor <- cor(w.go.df, method = 'spearman')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)

temp.cor2$key <- tstrsplit(temp.cor2$key, '_', keep = 1)[[1]]
temp.cor2$cluster2 <- tstrsplit(temp.cor2$cluster2, '_', keep = 1)[[1]]

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value, label = round(value,2))) +
  geom_tile() + # geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('Odds Ratio Correlation (Spearman)') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .25) +
  theme(axis.text.x = element_text(angle = 90))

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + # geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('Odds Ratio Correlation (Spearman)') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .1) +
  theme(axis.text.x = element_text(angle = 90))
```

### Signed FDR

```{r}
go.df$signed.FDR <- ifelse(go.df$Direction == 'down', -1*go.df$FDR, go.df$FDR)
out <- split(go.df, f = go.df$cluster)

# w.go.df

cnt = 1 

for (i in 1:length(out)){
  temp <- as.data.frame(out[[i]])
  rown <- as.character(temp$Name)
  coln <- paste0(names(out)[i],'_',colnames(temp)[11])
  colnames(temp)[11] <- coln
  temp <- data.frame(temp[,11])
  rownames(temp) <- rown
  temp <- data.frame(temp[go.bps,])
  colnames(temp) <- coln
  
  if(cnt == 1){
    w.go.df <- temp
    cnt = 2
  }
  else{
    w.go.df <- cbind(w.go.df, temp)
  }
  
}
```

```{r}
# summary(as.factor(go.df$cluster))
# length(go.bps)
```


```{r}
spearman.cor <- cor(w.go.df, method = 'spearman')

temp.cor <- as.data.frame(spearman.cor)

temp.cor2 <- gather(as.data.frame(temp.cor))

temp.cor2$cluster2 <- rep(rownames(temp.cor), length(rownames(temp.cor)))

# temp.cor2$value2 <- ifelse(temp.cor2$value == 1, 0, temp.cor2$value)
temp.cor2$key <- tstrsplit(temp.cor2$key, '_', keep = 1)[[1]]
temp.cor2$cluster2 <- tstrsplit(temp.cor2$cluster2, '_', keep = 1)[[1]]

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value, label = round(value,2))) +
  geom_tile() + # geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('Signed FDR Correlation (Spearman)') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .25) +
  theme(axis.text.x = element_text(angle = 90))

temp.cor2$value2 <- ifelse(temp.cor2$value ==1, 0, temp.cor2$value)

ggplot(temp.cor2, aes(x = key, y = cluster2, fill = value2, label = round(value,2))) +
  geom_tile() + # geom_text() + 
  xlab(NULL) + ylab(NULL) + ggtitle ('Signed FDR Correlation (Spearman)') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .05) +
  theme(axis.text.x = element_text(angle = 90))


```

### PCA

#### Odds Ratio

```{r}
cnt = 1 

for (i in 1:length(out)){
  temp <- as.data.frame(out[[i]])
  rown <- as.character(temp$Name)
  coln <- paste0(names(out)[i],'_',colnames(temp)[5])
  colnames(temp)[5] <- coln
  temp <- data.frame(temp[,5])
  rownames(temp) <- rown
  temp <- data.frame(temp[go.bps,])
  colnames(temp) <- coln
  
  if(cnt == 1){
    w.go.df <- temp
    cnt = 2
  }
  else{
    w.go.df <- cbind(w.go.df, temp)
  }
  
}
pca <- prcomp(w.go.df)

pca.sd <- pca$sdev
pca <- as.data.frame(pca$rotation)

pca.sd <- data.frame(pc = 1:9,
                     sd = pca.sd)

ggplot(pca.sd[1:9,], aes(x = pc, y = sd)) + geom_point() + geom_line() + theme_bw()
# colnames(pca) <- paste0(colnames(pca),' (',round(pca.sd,2),')')

pca$cluster <- tstrsplit(rownames(pca),'_', keep = 1)[[1]]
# pca

combos <- combn(paste0('PC',1:5),2)
# combos

# colnames(pca)[1:5] <- paste0(colnames(pca)[1:5], ' (SD = ', pca.sd[1:5], ')')

for (i in 1:dim(combos)[2]){
  # print(combos[1,i])
  temp.x <- combos[1,i]
  # print(combos[2,i])
  temp.y <- combos[2,i]
  
  print(ggplot(pca, aes_string(x = temp.x, y = temp.y, color = 'cluster')) + 
          geom_point() + scale_color_manual(values = color_pal2) +
          theme_bw() + ggtitle(paste0(temp.x, ' vs ', temp.y)))
}
```

#### Signed FDR

```{r}
go.df$signed.FDR <- ifelse(go.df$Direction == 'down', -1*go.df$FDR, go.df$FDR)
out <- split(go.df, f = go.df$cluster)

# w.go.df

cnt = 1 

for (i in 1:length(out)){
  temp <- as.data.frame(out[[i]])
  rown <- as.character(temp$Name)
  coln <- paste0(names(out)[i],'_',colnames(temp)[11])
  colnames(temp)[11] <- coln
  temp <- data.frame(temp[,11])
  rownames(temp) <- rown
  temp <- data.frame(temp[go.bps,])
  colnames(temp) <- coln
  
  if(cnt == 1){
    w.go.df <- temp
    cnt = 2
  }
  else{
    w.go.df <- cbind(w.go.df, temp)
  }
  
}
pca <- prcomp(w.go.df)

pca.sd <- pca$sdev
pca <- as.data.frame(pca$rotation)

pca.sd <- data.frame(pc = 1:9,
                     sd = pca.sd)

ggplot(pca.sd[1:9,], aes(x = pc, y = sd)) + geom_point() + geom_line() + theme_bw()
# colnames(pca) <- paste0(colnames(pca),' (',round(pca.sd,2),')')

pca$cluster <- tstrsplit(rownames(pca),'_', keep = 1)[[1]]
# pca

combos <- combn(paste0('PC',1:5),2)
# combos

# colnames(pca)[1:5] <- paste0(colnames(pca)[1:5], ' (SD = ', pca.sd[1:5], ')')

for (i in 1:dim(combos)[2]){
  # print(combos[1,i])
  temp.x <- combos[1,i]
  # print(combos[2,i])
  temp.y <- combos[2,i]
  
  print(ggplot(pca, aes_string(x = temp.x, y = temp.y, color = 'cluster')) + 
          geom_point() + scale_color_manual(values = color_pal2) +
          theme_bw() + ggtitle(paste0(temp.x, ' vs ', temp.y)))
}
```

## 5b Between Cluster GO Terms

### Generating the Data

```{r}
# de.genes <- data.frame(p_val = numeric(),
#                        avg_logFC = numeric(),
#                        pct.1 = numeric(),
#                        pct.2 = numeric(),
#                        p_val_adj = numeric(),
#                        cluster1 = character(),
#                        cluster2 = character(),
#                        gene = character())
# 
# # wbm@meta.data
# 
# combos <- as.data.frame(combn(unique(wbm$figure.labels3), 2))
# 
# mk.combos <- combos[,grepl('MK', combos)]
# non.mk.combos <- combos[,!grepl('MK',combos)]
# 
# for (i in 1:dim(mk.combos)[2]){
#   print(mk.combos[,i])
#   ct.1 <- mk.combos[1,i]
#   ct.2 <- mk.combos[2,i]
#   x <- FindMarkers(wbm,
#                    ident.1 = ct.1,
#                    ident.2 = ct.2,
#                    logfc.threshold = 0,
#                    min.pct = 0)
#   x$cluster1 <- ct.1
#   x$cluster2 <- ct.2
#   x$gene <- rownames(x)
#   x <- x[order(x$avg_log2FC, decreasing = T),]
#   de.genes <- rbind(de.genes,x)
# }
# 
# summary(as.factor(de.genes$cluster1))
# # table(rownames(de.genes))
# 
# write.table(de.genes, './data/v4/genes.for.LR.Path.between.MK.clusters.all.txt', sep = '\t')
# 
# de.genes <- data.frame(p_val = numeric(),
#                      avg_logFC = numeric(),
#                      pct.1 = numeric(),
#                      pct.2 = numeric(),
#                      p_val_adj = numeric(),
#                      cluster1 = character(),
#                      cluster2 = character(),
#                      gene = character())
# 
# ct.1 <- 'MkP'
# ct.2 <- 'MEP'
# 
# Idents(wbm) <- wbm$figure.labels3
# x <- FindMarkers(wbm, ident.1 = ct.1, ident.2 = ct.2,
#                  logfc.threshold = 0,
#                  min.pct = 0)
# x$cluster1 <- ct.1
# x$cluster2 <- ct.2
# x$gene <- rownames(x)
# x <- x[order(x$avg_log2FC, decreasing = T),]

# for (i in 1:dim(non.mk.combos)[2]){
#   print(non.mk.combos[,i])
#   print(paste0('Progress: ', i,' / ', dim(non.mk.combos)[2]))
#   ct.1 <- non.mk.combos[1,i]
#   ct.2 <- non.mk.combos[2,i]
#   x <- FindMarkers(wbm,
#                    ident.1 = ct.1,
#                    ident.2 = ct.2,
#                    logfc.threshold = 0,
#                    min.pct = 0)
  # x$cluster1 <- ct.1
  # x$cluster2 <- ct.2
  # x$gene <- rownames(x)
  # x <- x[order(x$avg_log2FC, decreasing = T),]
#   de.genes <- rbind(de.genes,x)
# }
```

### Exporting Results to LRPath Format

```{r}
# x <- read.table('./data/v4/genes.for.LR.Path.between.MK.clusters.all.txt', sep = '\t')
# 
# mus.ref <- data.table::fread('../mus_mus_gene_id.txt', header = T)
# colnames(mus.ref)[2] <- 'gene'
# 
# x$comp <- paste0(x$cluster1,'.vs.', x$cluster2)
# out <- split(x, f = x$comp)
# 
# for (i in 1:length(out)){
#   temp.name <- unique(out[[i]]$comp)
#   temp.name <- gsub('/','-',temp.name)
# 
#   print(temp.name)
# 
#   temp <- out[[i]]
#   print(summary(temp$gene %in% mus.ref$gene))
# 
#   temp.df <- merge(temp, mus.ref, by = 'gene')
# 
#   temp.df$entrez.name <- paste0('mmu:', temp.df$GeneID)
# 
#   temp.df <- temp.df[,c('GeneID','p_val_adj','avg_log2FC')]
# 
#   write.table(temp.df, paste0('./data/v4/LRPath.gene.between.clusters.lists/',
#                               temp.name,'.txt'),
#               sep = '\t', col.names = F, quote = F, row.names = F)
# }
```

## Reading in all the Terms

```{r}
# setwd('data/v4/LRPath.results/')
files = list.files('./data/v4/LRPath.between.clusters.results/')
files <- files[grepl('.txt',files)]
# files

cnt <- 1

for (i in files){
  # print(i)
  x <- fread(paste0('./data/v4/LRPath.between.clusters.results/',i))
  x$cluster <- tstrsplit(i,'.txt', keep = 1)[[1]]
  
  if (cnt == 1){
    go.df <- x
    cnt <- 2
  }
  else{
    go.df <- rbind(go.df, x)
  }
}

# summary(as.factor(go.df$cluster))
# length(unique(go.df$cluster))
# summary(as.factor(go.df$ConceptType))
```

## Subsetting to just GOBP

```{r}
go.df <- go.df[go.df$ConceptType == 'GOBP',]

tmp <- table(go.df$Name, go.df$cluster)

# summary(as.factor(go.df$Name))

go.bps <- rownames(tmp[rowSums(tmp) == 19,])

# length(go.bps)
```


## 5c Meta-Analysis of all GO Terms


# 6 Plots for all Clusters

All the Violin Plots

## 6a BoxPlots

### nCount_RNA

```{r}
# head(wbm@meta.data)
# ggplot(wbm@meta.data, aes(x = figure.labels3, y = log10(nCount_RNA), fill = Condition3)) + geom_violin()

ggplot(wbm@meta.data, aes(x = figure.labels3, y = log10(nCount_RNA))) + 
  geom_boxplot() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

ggplot(wbm@meta.data, aes(x = figure.labels3, y = log10(nCount_RNA), 
                          fill = Condition3)) + 
  geom_boxplot() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

``` 

### nFeature_RNA

```{r}
ggplot(wbm@meta.data, aes(x = figure.labels3, y = log10(nFeature_RNA))) + 
  geom_boxplot() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

ggplot(wbm@meta.data, aes(x = figure.labels3, y = log10(nFeature_RNA), 
                          fill = Condition3)) + 
  geom_boxplot() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

``` 

### percent.mt

```{r}
ggplot(wbm@meta.data, aes(x = figure.labels3, y = percent.mt)) + 
  geom_boxplot() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

ggplot(wbm@meta.data, aes(x = figure.labels3, y = percent.mt, 
                          fill = Condition3)) + 
  geom_boxplot() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

``` 

## 6b Interaction Plots

### nCount & nFeature

```{r}
ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = log10(nFeature_RNA))) +
  geom_point() + theme_bw() + geom_smooth()

ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = log10(nFeature_RNA),
                          colour = Condition3)) +
  scale_color_manual(values = c('black','red','limegreen','blue')) +
  geom_point() + theme_bw() + geom_smooth(se = F)

ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = log10(nFeature_RNA),
                          colour = figure.labels3)) +
  scale_color_manual(values = color_pal2) +
  geom_point() + theme_bw() 

for (i in unique(wbm$figure.labels3)){
        # print(i)
        
        wbm$highlight.cells <- ifelse(wbm$figure.labels3 == i, i,'All Others')
        
        print(ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = log10(nFeature_RNA),
                          colour = highlight.cells)) +
          scale_color_manual(values = c('lightgrey','blue')) +
          geom_point(aes(alpha = 0.2)) + theme_bw() + geom_smooth())
        
}
```

### nCount & percent.mt

```{r}
ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = percent.mt)) +
  geom_point() + theme_bw() + geom_smooth()

ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = percent.mt,
                          colour = Condition3)) +
  scale_color_manual(values = c('black','red','limegreen','blue')) +
  geom_point() + theme_bw() + geom_smooth(se = F)

ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = percent.mt,
                          colour = figure.labels3)) +
  scale_color_manual(values = color_pal2) +
  geom_point() + theme_bw() 

for (i in unique(wbm$figure.labels3)){
        # print(i)
        
        wbm$highlight.cells <- ifelse(wbm$figure.labels3 == i, i,'All Others')
        
        print(ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = percent.mt,
                          colour = highlight.cells)) +
          scale_color_manual(values = c('lightgrey','blue')) +
          geom_point(aes(alpha = 0.2)) + theme_bw() + geom_smooth())
        
}
# wbm$b.cells <- ifelse(wbm$figure.labels3 == 'B-cell', 'B-cell', 'All Others')
# 
# ggplot(wbm@meta.data, aes(x = log10(nCount_RNA), y = percent.mt,
#                           colour = b.cells)) +
#   scale_color_manual(values = c('lightgrey','blue')) +
#   geom_point(aes(alpha = 0.2)) + theme_bw() + geom_smooth()
```

### nFeature & percent.mt

```{r}
ggplot(wbm@meta.data, aes(x = log10(nFeature_RNA), y = percent.mt)) +
  geom_point() + theme_bw() + geom_smooth()

ggplot(wbm@meta.data, aes(x = log10(nFeature_RNA), y = percent.mt,
                          colour = Condition3)) +
  scale_color_manual(values = c('black','red','limegreen','blue')) +
  geom_point() + theme_bw() + geom_smooth(se = F)

ggplot(wbm@meta.data, aes(x = log10(nFeature_RNA), y = percent.mt,
                          colour = figure.labels3)) +
  scale_color_manual(values = color_pal2) +
  geom_point() + theme_bw() 

for (i in unique(wbm$figure.labels3)){
        # print(i)
        
        wbm$highlight.cells <- ifelse(wbm$figure.labels3 == i, i,'All Others')
        
        print(ggplot(wbm@meta.data, aes(x = log10(nFeature_RNA), y = percent.mt,
                          colour = highlight.cells)) +
          scale_color_manual(values = c('lightgrey','blue')) +
          geom_point(aes(alpha = 0.2)) + theme_bw() + geom_smooth())
        
}
# 
# ggplot(wbm@meta.data, aes(x = log10(nFeature_RNA), y = percent.mt,
#                           colour = b.cells)) +
#   scale_color_manual(values = c('lightgrey','blue')) +
#   geom_point(aes(alpha = 0.2)) + theme_bw() + geom_smooth()
```




# 7 Trajectory

Looking at the trajectory analysis for just the MK cell types (MK, MkP, MEP). 

```{r}
library(monocle)

x <- subset(wbm, figure.labels3 %in% c('MK-1','MK-2','MEP','MkP'))

data <- as(as.matrix(x@assays$RNA@data), 'sparseMatrix')


pd <- new('AnnotatedDataFrame', data = x@meta.data)
fData <- data.frame(gene_short_name = row.names(data), row.names = row.names(data))
fd <- new('AnnotatedDataFrame', data = fData)

cds <- newCellDataSet(data,
                              phenoData = pd,
                              featureData = fd,
                              lowerDetectionLimit = 0.5,
                              expressionFamily = negbinomial.size())


cds <- estimateSizeFactors(cds)

cds <- estimateDispersions(cds)

cds <- detectGenes(cds, min_expr = 0.1)

expressed_genes <- rownames(subset(fData(cds), num_cells_expressed >= 10))

fData(cds)$use_for_ordering <- fData(cds)$num_cells_expressed > 0.05 * ncol (cds)

cds <- reduceDimension(cds, reduction_method = 'DDRTree')

cds <- orderCells(cds)

plot_cell_trajectory(cds, color_by = 'figure.labels3')
?plot_cell_trajectory
plot_cell_trajectory(cds, color_by = 'Pseudotime')
plot_cell_trajectory(cds, color_by = 'State')

table(cds$State, cds$figure.labels3)
```


Should maybe also show it including Erythroid and/or CMP.

# 8 Subclustering of MK Lineages

Looking at subclustering between MEP, MkP, and MK.

```{r}
x <- subset(wbm, figure.labels3 %in% c('MK-1','MK-2','MEP','MkP'))

mk <- NormalizeData(x, normalization.method = 'LogNormalize', scale.factor = 10000)

mk <- FindVariableFeatures(mk, selection.method = 'vst', nfeatures = 2000)

summary(VariableFeatures(mk) %in% VariableFeatures(wbm))

mk <- ScaleData(mk, features = rownames(mk))

mk <- RunPCA(mk, features = VariableFeatures(mk), verbose = F)

ElbowPlot(mk) # going with 10

mk <- FindNeighbors(mk, dims = 1:10)

for (i in seq(0,1,.1)){
  mk <- FindClusters(mk, resolution = i)
}

mk <- RunUMAP(mk, dims = 1:10)

mk <- RunTSNE(mk, dims = 1:10)

# res.columns <- colnames(mk@meta.data)[grepl('RNA_snn',
#                                             colnames(mk@meta.data), ignore.case = T)]
# 
# for (i in res.columns){
#   # print(i)
#   print(DimPlot(mk, group.by = i, reduction = 'umap', label = T, repel = T) + NoLegend())
# }

Idents(mk) <- mk$RNA_snn_res.0.4

mk$mk.clusters <- mk$RNA_snn_res.0.4

```

## UMAP 

```{r}
DimPlot(mk, label = T, repel = T)
DimPlot(mk, label = T, repel = T, split.by = 'Condition3', ncol = 2) +
  theme_bw()


DimPlot(mk, label = T, repel = T, group.by = 'figure.labels3') + ggtitle(NULL)
DimPlot(mk, label = T, repel = T, group.by = 'figure.labels3', 
        split.by = 'Condition3', ncol = 2) + theme_bw() + ggtitle(element_blank())

temp.tbl <- as.data.frame(table(Idents(mk), mk$figure.labels3))

temp.tbl$var2.total <- NA
for (i in unique(temp.tbl$Var2)){
  temp.tbl[temp.tbl$Var2 == i,]$var2.total <- sum(temp.tbl[temp.tbl$Var2 == i,]$Freq)
}

temp.tbl$var1.total <- NA
for (i in unique(temp.tbl$Var1)){
  temp.tbl[temp.tbl$Var1 == i,]$var1.total <- sum(temp.tbl[temp.tbl$Var1 == i,]$Freq)
}

temp.tbl$pct.var1 <- temp.tbl$Freq / temp.tbl$var1.total
temp.tbl$pct.var2 <- temp.tbl$Freq / temp.tbl$var2.total

ggplot(temp.tbl, aes(x = Var1, y = Var2, fill = pct.var1, label = Freq)) + 
  geom_tile() + 
  theme_bw() +
  geom_text(size = 8) + xlab('Subclusters') + ylab('Original Cluster') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(temp.tbl, aes(x = Var1, y = Var2, fill = pct.var2, label = Freq)) + 
  geom_tile() + 
  theme_bw() +
  geom_text(size = 8) + xlab('Subclusters') + ylab('Original Cluster') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```

```{r}
mk$combo.label <- paste0(mk$Experiment,'_', mk$figure.labels3)

temp.tbl <- as.data.frame(table(Idents(mk), mk$combo.label))

temp.tbl$var2.total <- NA
for (i in unique(temp.tbl$Var2)){
  temp.tbl[temp.tbl$Var2 == i,]$var2.total <- sum(temp.tbl[temp.tbl$Var2 == i,]$Freq)
}

temp.tbl$var1.total <- NA
for (i in unique(temp.tbl$Var1)){
  temp.tbl[temp.tbl$Var1 == i,]$var1.total <- sum(temp.tbl[temp.tbl$Var1 == i,]$Freq)
}

temp.tbl$pct.var1 <- temp.tbl$Freq / temp.tbl$var1.total
temp.tbl$pct.var2 <- temp.tbl$Freq / temp.tbl$var2.total

ggplot(temp.tbl, aes(x = Var1, y = Var2, fill = pct.var1, label = Freq)) + 
  geom_tile() + 
  theme_bw() +
  geom_text(size = 8) + xlab('Subclusters') + ylab('Original Cluster') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)

ggplot(temp.tbl, aes(x = Var1, y = Var2, fill = pct.var2, label = Freq)) + 
  geom_tile() + 
  theme_bw() +
  geom_text(size = 8) + xlab('Subclusters') + ylab('Original Cluster') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5)
```



## tSNE 

```{r}
DimPlot(mk, label = T, repel = T, reduction = 'tsne')
DimPlot(mk, label = T, repel = T, split.by = 'Condition3', ncol = 2, 
        reduction = 'tsne') +
  theme_bw()


DimPlot(mk, label = T, repel = T, group.by = 'figure.labels3', 
        reduction = 'tsne')
DimPlot(mk, label = T, repel = T, group.by = 'figure.labels3', 
        split.by = 'Condition3', ncol = 2, reduction = 'tsne') + 
  theme_bw()
```

## PC Plots

```{r}
combos <- combn(1:5,2)

for (i in 1:dim(combos)[2]){
  # print(i)
  print(DimPlot(mk, reduction = 'pca', 
                dims = combos[,i]))
}

for (i in 1:dim(combos)[2]){
  # print(i)
  print(DimPlot(mk, reduction = 'pca', 
                dims = combos[,i], group.by = 'Condition3'))
}

for (i in 1:dim(combos)[2]){
  # print(i)
  print(DimPlot(mk, reduction = 'pca', 
                dims = combos[,i], group.by = 'figure.labels3'))
}
```

Saving the metadata for velocity analysis


```{r}
# mk@reductions$umap@cell.embeddings[,'UMAP_1']
mk$UMAP1 <- mk@reductions$umap@cell.embeddings[,'UMAP_1']
mk$UMAP2 <- mk@reductions$umap@cell.embeddings[,'UMAP_2']
mk$tsne1 <- mk@reductions$tsne@cell.embeddings[,'tSNE_1']
mk$tsne2 <- mk@reductions$tsne@cell.embeddings[,'tSNE_2']

write.table(mk@meta.data, './data/v4/seurat.MKLineage.metadata.txt', sep = '\t', quote = F)
```




# Inflammatory Score

Either using the list previously published, the list added onto by Priya, or also a list from GO term analysis.

# Velocity Analysis

Using scVelo on the cluster

# 9 Erythrocyte Naming

## Unique Markers

```{r}
Idents(wbm) <- wbm$figure.labels3
ery.markers <- FindMarkers(wbm,
            ident.1 = 'Erythroid',
            only.pos = T)

# head(ery.markers)
ery.markers$pct.diff <- ery.markers$pct.1 - ery.markers$pct.2
head(ery.markers)
```

# 10 MK-1 Doublets?

```{r}
DimPlot(wbm, group.by = 'seurat_clusters', label= T, repel = T)
wbm$mk.fine.labels <- paste0(wbm$figure.labels3, '_', wbm$seurat_clusters)
DimPlot(wbm, group.by = 'mk.fine.labels', label = T, repel = T)

```

## General Markers

```{r}
all.gran.markers <- FindMarkers(wbm, ident.1 = 'Granulocyte', only.pos = T)
# head(all.gran.markers)
gran.0.1.markers <- FindMarkers(wbm, ident.1 = c('Granulocyte_0','Granulocyte_1'),
                                group.by = 'mk.fine.labels', only.pos = T)
mk2.markers <- FindMarkers(wbm, ident.1 = 'MK-2', only.pos = T)
mk1.markers <- FindMarkers(wbm, ident.1 = 'MK-1', only.pos = T)
```

```{r}
# head(all.gran.markers)

for (i in rownames(all.gran.markers)[1:6]){
  print(VlnPlot(wbm, features = i) + ggtitle(paste0('Gran Marker ',i)))
}
```

```{r}
# head(gran.0.1.markers)

for (i in rownames(gran.0.1.markers)[1:6]){
  print(VlnPlot(wbm, features = i) + ggtitle(paste0('Gran Marker ',i)))
}
```

```{r}
# head(mk2.markers)

for (i in rownames(mk2.markers)[1:6]){
  print(VlnPlot(wbm, features = i) + ggtitle(paste0('MK-2 Marker ',i)))
}
```

```{r}
# head(mk1.markers)

for (i in rownames(mk1.markers)[1:6]){
  print(VlnPlot(wbm, features = i) + ggtitle(paste0('MK-1 Marker ',i)))
}
```

## Comparing Populations

### MK-1 vs Granulocytes

```{r}
gran.vs.mk1 <- FindMarkers(wbm, ident.1 = 'MK-1', 
                           ident.2 = 'Granulocyte')

gran.vs.mk1 <- gran.vs.mk1[gran.vs.mk1$p_val < 0.05,]
gran.vs.mk1 <- gran.vs.mk1[order(gran.vs.mk1$avg_log2FC, decreasing = T),]

# head(gran.vs.mk1)

for (i in rownames(gran.vs.mk1)[1:6]){
  print(VlnPlot(wbm, features = i) + ggtitle(paste0('MK-1 vs Gran ',i)))
}

# tail(gran.vs.mk1)

for (i in rownames(tail(mk1.markers))){
  print(VlnPlot(wbm, features = i) + ggtitle(paste0('MK-1 Marker ',i)))
}
```

### MK-1 vs Specific Granulocytes

```{r}
gran.vs.mk1 <- FindMarkers(wbm, ident.1 = 'MK-1_14', 
                           ident.2 = c('Granulocyte_0', 'Granulocyte_1'),
                           group.by = 'mk.fine.labels')

gran.vs.mk1 <- gran.vs.mk1[gran.vs.mk1$p_val < 0.05,]
gran.vs.mk1 <- gran.vs.mk1[order(gran.vs.mk1$avg_log2FC, decreasing = T),]

# head(gran.vs.mk1)

for (i in rownames(gran.vs.mk1)[1:6]){
  print(VlnPlot(wbm, features = i, group.by = 'mk.fine.labels') + 
          ggtitle(paste0('MK-1  ',i)))
}

# tail(gran.vs.mk1)

for (i in rownames(tail(mk1.markers))){
  print(VlnPlot(wbm, features = i, group.by = 'mk.fine.labels') +
          ggtitle(paste0('Gran 0 and 1 Marker ',i)))
}
```

# 11 SCENIC analysis

# 12 CD41 CITE-seq

CD41 (Itga2b) data was collected during the sequencing process. The enriched samples, were enriched for CD41, so we should see that.

The cite-seq analysis was performed in HTO.Rmd


```{r}
cite.seq <- read.table('./data/htos.for.jun.txt', sep = '\t')

cite.seq <- as.data.frame(t(cite.seq))

summary(as.numeric(cite.seq$CD41))

c.o.i <- colnames(cite.seq)[-7]

for (i in 1:length(c.o.i)){
        # print(c.o.i[i])
        cite.seq[,i +7 ] <- NA
        cite.seq[,i + 7] <- log10(as.numeric(cite.seq[,i]))
}

colnames(cite.seq)[8:13] <- paste0(colnames(cite.seq)[1:6],'.log')

ggplot(cite.seq, aes(x = CD41.log)) + geom_density() + theme_bw() +
        geom_vline(xintercept = c(1.5,2,2.5), colour = 'red', linetype = 2)
```

Looking at different cutoffs to label "high CD41". Originally hoping it would be a bimodal distribution, but since it is unimodal need to figure how to ID "high" CD41 cells.

Going to go with 10**1.5 for now

```{r}
cite.seq$cd41.high <- ifelse(cite.seq$CD41.log > 1.5, T, F)

summary(cite.seq$cd41.high)
```

## Merging with Seurat Object

```{r}
summary(colnames(wbm) %in% paste0(rownames(cite.seq),'-1'))

wbm@meta.data[!(colnames(wbm) %in% paste0(rownames(cite.seq),'-1')),]
```

Oddly enough there is one row in the Seurat data that doesn't have cite-seq information...a bit odd

## 12a Quantification of CD41 "High"

### CD41 in UMAP

```{r}
rownames(cite.seq) <- paste0(rownames(cite.seq),'-1')

cite.seq <- cite.seq[colnames(wbm),]

wbm@meta.data$cd41.log <- cite.seq$CD41.log

wbm@meta.data$cd41.high <- cite.seq$cd41.high

FeaturePlot(wbm, features = 'cd41.log')

FeaturePlot(wbm, features = 'cd41.high')
```

### BoxPlots by Cluster

```{r}
ggplot(wbm@meta.data, aes(x = figure.labels3, y = cd41.log)) + 
  geom_boxplot() + 
  theme_bw() +
  geom_hline(yintercept = 1.5, linetype = 2, colour = 'black') +
  theme(axis.text.x = element_text(angle = 90))

ggplot(wbm@meta.data, aes(x = figure.labels3, y = cd41.log, 
                          color = Condition3)) + 
  geom_boxplot() + 
  geom_hline(yintercept = 1.5, linetype = 2, colour = 'black') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

```

### Heatmap by Cluster

```{r}
tmp <- as.data.frame(table(wbm$cd41.high, wbm$figure.labels3))
tmp$Var1 <- ifelse(tmp$Var1 == T, 'High','Low')
tmp$var1.sum <- NA

for (i in unique(tmp$Var1)){
  tmp[tmp$Var1 == i,]$var1.sum <- sum(tmp[tmp$Var1 ==i,]$Freq)
}

tmp$var2.sum <- NA

for (i in unique(tmp$Var2)){
  tmp[tmp$Var2 == i,]$var2.sum <- sum(tmp[tmp$Var2 ==i,]$Freq)
}

tmp$var1.perc <- tmp$Freq / tmp$var1.sum

tmp$var2.perc <- tmp$Freq / tmp$var2.sum

ggplot(tmp, aes(x = Var1, y = Var2, fill = var1.perc)) +
  geom_tile() + geom_text(aes(label = Freq)) +
  xlab('CD41 Level') + ylab('CellType') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')

ggplot(tmp, aes(x = Var1, y = Var2, fill = var2.perc)) +
  geom_tile() + geom_text(aes(label = Freq)) +
  xlab('CD41 Level') + ylab('CellType') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', 
                       midpoint = 0.5)
```

This is what we would expect. MK-1, MK-2 and MkP are the celltypes with the highest proportion of cells that are CD41 high. We see that in the boxplots and heatmap.

### BoxPLots by Sample

```{r}

ggplot(wbm@meta.data, aes(y = Condition3, x = cd41.log)) + 
  geom_boxplot() + 
  geom_vline(xintercept = 1.5, linetype = 2, colour = 'red') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
tmp <- as.data.frame(table(wbm$cd41.high, wbm$Condition3))
tmp$Var1 <- ifelse(tmp$Var1 == T, 'High','Low')
tmp$var1.sum <- NA

for (i in unique(tmp$Var1)){
  tmp[tmp$Var1 == i,]$var1.sum <- sum(tmp[tmp$Var1 ==i,]$Freq)
}

tmp$var2.sum <- NA

for (i in unique(tmp$Var2)){
  tmp[tmp$Var2 == i,]$var2.sum <- sum(tmp[tmp$Var2 ==i,]$Freq)
}

tmp$var1.perc <- tmp$Freq / tmp$var1.sum

tmp$var2.perc <- tmp$Freq / tmp$var2.sum

ggplot(tmp, aes(x = Var1, y = Var2, fill = var1.perc)) +
  geom_tile() + geom_text(aes(label = Freq)) +
  xlab('CD41 Level') + ylab('Sample') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')

ggplot(tmp, aes(x = Var1, y = Var2, fill = var2.perc)) +
  geom_tile() + geom_text(aes(label = Freq)) +
  xlab('CD41 Level') + ylab('Sample') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', 
                       midpoint = 0.5)

ggplot(tmp, aes(x = Var1, y = Var2, fill = var2.perc)) +
  geom_tile() + geom_text(aes(label = round(var2.perc,2))) +
  xlab('CD41 Level') + ylab('Sample') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', 
                       midpoint = 0.5)
```

Higher percentage of high CD41 levels in the enr. samples.


## 12b Plots between Itga2b and CD41

```{r}
cnts <- wbm@assays$RNA@scale.data

wbm$itga2b.levels <- cnts['Itga2b',]

wbm$cd41 <- cite.seq$CD41

wm <- wbm@meta.data

wm <- wm[!is.na(wm$cd41.log),]

itga2b.vs.cd41 <- lm(itga2b.levels ~ cd41, data = wm)

```

### Using all Cells

```{r}
lm(cd41 ~ itga2b.levels, data = wm)

wm$cd41 <- as.numeric(wm$cd41)

cr <- cor(wm$itga2b.levels, wm$cd41)

ggplot(wm, aes(x = itga2b.levels, y = cd41)) + geom_point() +
  geom_smooth()

ggplot(wm, aes(x = itga2b.levels, y = cd41.log)) + geom_point() +
  geom_smooth()
```

### Using cells expressing Itga2b

```{r}
wm2 <- wm[wm$itga2b.levels != min(wm$itga2b.levels),]

lm(cd41 ~ itga2b.levels, data = wm2)

cor(wm2$cd41, wm2$itga2b.levels)

cor(wm2$cd41.log, wm2$itga2b.levels)

ggplot(wm2, aes(x = itga2b.levels, y = cd41)) + geom_point() +
  geom_smooth()

ggplot(wm2, aes(x = itga2b.levels, y = cd41.log)) + geom_point() +
  geom_smooth()
```

## 12c Cross Tab w/ binary Itga2b

```{r}
ggplot(wbm@meta.data, aes(x = itga2b.levels)) + geom_density()
ggplot(wbm@meta.data, aes(x = log(itga2b.levels))) + geom_density()

p <- as.data.frame(wbm@assays$RNA@counts['Itga2b',])
colnames(p) <- 'counts'
ggplot(p, aes(x = log(counts))) + geom_density() +
  geom_vline(xintercept = .4, colour = 'red', linetype = 'dashed')

wbm$itga2b.high <- ifelse(log(wbm$itga2b.levels) > .4,T, F)
wbm$itga2b.high <- ifelse(is.na(wbm$itga2b.high), F,T)
summary(wbm$itga2b.high)

ggplot(wbm@meta.data, aes(x = itga2b.high, y = itga2b.levels)) + 
  geom_boxplot()

# ggplot(wbm@meta.data, aes(color = itga2b.high, y = itga2b.levels)) + 
#   geom_density()

```

Setting the cutoff for itga2b high at .4 for the log(counts of Itga2b)

### Cluster vs Status

```{r}
table(wbm$itga2b.high, wbm$cd41.high)

wbm$itga2b.cd41.status <- ifelse(wbm$itga2b.high == T & wbm$cd41.high == T, 'Both',
                                 ifelse(wbm$itga2b.high == T, 'Itga2b only',
                                        ifelse(wbm$cd41.high == T, 'CD41 only',
                                               'Neither')))

DimPlot(wbm, group.by = 'itga2b.cd41.status', cols = c('black', 'red','blue',
                                                       'lightgrey'))

table(wbm$figure.labels3, wbm$itga2b.cd41.status)
```

```{r}
tmp <- as.data.frame(table(wbm$figure.labels3, wbm$itga2b.cd41.status))

tmp$Var1.sum <- NA
for (i in unique(tmp$Var1)){
  tmp[tmp$Var1 ==i,]$Var1.sum <- sum(tmp[tmp$Var1 == i,]$Freq)
}

tmp$Var2.sum <- NA
for (i in unique(tmp$Var2)){
  tmp[tmp$Var2 ==i,]$Var2.sum <- sum(tmp[tmp$Var2 == i,]$Freq)
}
tmp$var1.perc <- tmp$Freq / tmp$Var1.sum
tmp$var2.perc <- tmp$Freq / tmp$Var2.sum

ggplot(tmp, aes(x = Var2, y = Var1, fill = var1.perc, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text() +
  scale_fill_gradient2(low = 'blue', mid = 'white' , high = 'red', 
                       midpoint = 0.5)

ggplot(tmp, aes(x = Var2, y = Var1, fill = var2.perc, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text() +
  scale_fill_gradient2(low = 'blue', mid = 'white' , high = 'red', 
                       midpoint = 0.5)
```

### Sample vs Status

```{r}
DimPlot(wbm, split.by = 'Condition3', ncol = 2,
        group.by = 'itga2b.cd41.status', cols = c('black', 'red','blue',
                                                       'lightgrey')) +
  theme_bw()

table(wbm$Condition3, wbm$itga2b.cd41.status)
```

```{r}
tmp <- as.data.frame(table(wbm$Condition3, wbm$itga2b.cd41.status))

tmp$Var1.sum <- NA
for (i in unique(tmp$Var1)){
  tmp[tmp$Var1 ==i,]$Var1.sum <- sum(tmp[tmp$Var1 == i,]$Freq)
}

tmp$Var2.sum <- NA
for (i in unique(tmp$Var2)){
  tmp[tmp$Var2 ==i,]$Var2.sum <- sum(tmp[tmp$Var2 == i,]$Freq)
}
tmp$var1.perc <- tmp$Freq / tmp$Var1.sum
tmp$var2.perc <- tmp$Freq / tmp$Var2.sum

ggplot(tmp, aes(x = Var2, y = Var1, fill = var1.perc, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text() +
  scale_fill_gradient2(low = 'blue', mid = 'white' , high = 'red', 
                       midpoint = 0.5)

ggplot(tmp, aes(x = Var2, y = Var1, fill = var2.perc, label = Freq)) +
  geom_tile() + theme_bw() +
  geom_text() +
  scale_fill_gradient2(low = 'blue', mid = 'white' , high = 'red', 
                       midpoint = 0.5)

mid <- mean(tmp$var1.perc)

ggplot(tmp, aes(x = Var2, y = Var1, fill = var1.perc, label = round(var1.perc,2))) +
  geom_tile() + theme_bw() +
  geom_text() +
  scale_fill_gradient2(low = 'blue', mid = 'white' , high = 'red',
                       midpoint = mid)

mid <- mean(tmp$var2.perc)

ggplot(tmp, aes(x = Var2, y = Var1, fill = var2.perc, label = round(var2.perc,2))) +
  geom_tile() + theme_bw() +
  geom_text() +
  scale_fill_gradient2(low = 'blue', mid = 'white' , high = 'red',
                       midpoint = mid)
```