---
title: "Monocle Trajectories"
author: "D. Ford Hannum"
date: "6/17/2020"
output: 
        html_document:
                toc: true
                toc_depth: 3
                number_sections: false
                theme: united
                highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(Seurat)
library(ggplot2)
library(data.table)
library(monocle)
```

Looking at monocle trajectories for the scWBM project, experiment 1. Going to look at three different trajectories:

* all data
* MKs, erythrocytes and MEPs
* myeloid
* myeloblast (granulocytes, monocytes, and macrophages)


```{r loading data}
# NOT SURE WHY THE IMPORT DOESN'T WORK, JUST GOING TO IMPORT THE RAW DATA INTO MONOCLE
# ?importCDS()
# seuratWBM <- readRDS('./data/wbm_clustered_filtered_named.rds')
# x <- importCDS(seuratWBM)

mtx <- readMM('./data/filtered_feature_bc_matrix/matrix.mtx')
features <- read.table('./data/filtered_feature_bc_matrix/features.tsv')
colnames(features)[2] <- 'gene_short_name'

barcodes <- read.table('./data/filtered_feature_bc_matrix/barcodes.tsv')

pd <- new("AnnotatedDataFrame", data = barcodes)
fd <- new("AnnotatedDataFrame", data = features)

library(Matrix)

cds <- newCellDataSet(as(as.matrix(mtx),"sparseMatrix"),
                      phenoData = pd,
                      featureData = fd,
                      lowerDetectionLimit = 0.5,
                      expressionFamily = negbinomial.size())

# HAVING ISSUES RUNNING estimateDispresions SO GOING TO FILTER TO THE CELLS AND GENES FROM THE SEURAT ANALYSIS

swbm <- readRDS('./data/wbm_clustered_filtered_named.rds')
valid_cells <- colnames(swbm)

# there is a -1 in front of the monocle cells

#summary(as.factor(tstrsplit(pData(cds)$V1, '-', keep = 2)[[1]]))

valid_cells <- paste0(valid_cells,'-1')

valid_cells <- rownames(subset(pData(cds), 
                               V1 %in% valid_cells))
#head(valid_cells)

cds <- cds[,valid_cells]
```

```{r doing all cells}
cds <- estimateSizeFactors(cds)
cds <- estimateDispersions(cds)

cds <- detectGenes(cds, min_expr = 0.1)

#summary(fData(cds)$num_cells_expressed)

expressed_genes <- rownames(subset(fData(cds), num_cells_expressed >= 10))
```


```{r adding info to the pData from Seurat}
# DIDN'T WANT TO MESS UP THE MONOCLE OBJECT AND START OVER, SO cds1 IS A BACKUP
#cds1 <- cds
#summary(pData(cds)$V1 == paste0(rownames(swbm@meta.data),'-1'))
pData(cds)[,c('nCountRNA', 'nFeature_RNA', 'celltype','condition', 'percent.mt',
              'seurat_clusters', 'cluster_IDs')] <- swbm@meta.data[,c(2,3,5,6,8,11,15)]
colnames(pData(cds))[1] <- 'cell'
#head(pData(cds))
```

# All Data

```{r ordering genes}
cds <- detectGenes(cds, min_expr = 0.1)

fData(cds)$use_for_ordering <- fData(cds)$num_cells_expressed > 0.05 * ncol(cds)

# THE BELOW CODE TOOK FOREVER TO RUN SO I STOPPED IT
#plot_pc_variance_explained(cds, return_all = F, max_components = 10)

cds <- reduceDimension(cds,
                       max_components = 2,
                       norm_method = 'log',
                       num_dim = 3,
                       reduction_method = 'tSNE',
                       verbose = T)

cds <- clusterCells(cds, verbose = F)

# checking clustering results

# not sure what to read into these figures
plot_cell_clusters(cds, color_by = 'Cluster')
plot_cell_clusters(cds, color_by = 'cluster_IDs')
plot_cell_clusters(cds, color_by = 'condition')

plot_rho_delta(cds, rho_threshold = 2, delta_threshold = 4)
cds <- clusterCells(cds,
                    rho_threshold = 2,
                    delta_threshold = 4,
                    sskip_rho_sigma = T,
                    verbose = F)

plot_cell_clusters(cds, color_by = 'Cluster')
plot_cell_clusters(cds, color_by = 'condition')
plot_cell_clusters(cds, color_by = 'cluster_IDs')

clust_DEG_genes <- differentialGeneTest(cds[expressed_genes,],
                                        fullModelFormulaStr = "~Cluster",
                                        cores = 1)

cds_ordering_genes <- row.names(clust_DEG_genes)[order(clust_DEG_genes$qval)][1:1000]
cds <- setOrderingFilter(cds,
                         ordering_genes = cds_ordering_genes)
cds <- reduceDimension(cds, method = 'DDRTree')
cds <- orderCells(cds)

plot_cell_trajectory(cds, color_by = "cluster_IDs")

#summary(as.factor(pData(cds)$cluster_IDs))

pData(cds)$cell_type_broad <- ifelse(pData(cds)$cluster_IDs %in% c(paste0("Granulocyte",1:4), 'Stem Cell'), 'Granulocyte',
                                     ifelse(pData(cds)$cluster_IDs %in% c('Monocyte','Macrophage'), 'Mono/Macro',
                                            ifelse(pData(cds)$cluster_IDs %in% c('MK','Erythroid','MEP'), 'MK/Ery/MEP',
                                                   'Lymphocyte')))
table(pData(cds)$cluster_IDs, pData(cds)$cell_type_broad)

# pData(cds)$pot_progenitors <- ifelse(cds$cluster_IDs %in% c('MEP', 'B-cell pro','Stem Cell'), 'Potential Progenitors',
#                                      'Further Cell States')

cds$prog <- ifelse(cds$cluster_IDs %in% c('MEP', 'B-cell pro','Stem Cell'), as.character(cds$cluster_IDs), 'Further States')

cds$type <- ifelse(cds$cluster_IDs %in% c('MEP', 'Erythroid','MEP'), 'MK/Ery/MEP',
                   ifelse(cds$cluster_IDs %in% c('B cell', 'B-cell pro','T-cell/NK'), 'Lymphocyte','Myeloid'))

plot_cell_trajectory(cds, color_by = "cell_type_broad")
#plot_cell_trajectory(cds, color_by = 'pot_progenitors')
plot_cell_trajectory(cds, color_by = 'prog')
plot_cell_trajectory(cds, color_by = 'condition')
plot_cell_trajectory(cds, color_by = 'type')
```

This is the cell trajectory using unsupervised methods. Looks pretty good overall, with the exception of monocytes and macrophages being at the end of the MK/Ery/MEP trajectory. Had low expectations for using all these cells. The clusters we had labeled in Seurat as progenitor populations had differing results. The stem cells, which we've decided is a poor name as in more likely a granulocyte progenitor/neutrophil, showed close to the base, whereas MEPs were disperesed and B-cell prog were actually at the end of the trajectory. Would be more useful to look at the separated populations, which is what will come next. 


# MK, Erythroid and MEP Trajectory

```{r mep trajectory}
Mcds <- cds[,cds$cluster_IDs %in% c('MEP',"MK", "Erythroid")]
#dim(Mcds)

Mcds <- estimateSizeFactors(Mcds)
Mcds <- estimateDispersions(Mcds)

Mcds <- detectGenes(Mcds, min_expr = 0.1)

fData(Mcds)$use_for_ordering <- fData(Mcds)$num_cells_expressed > 0.05 * ncol(Mcds)

Mcds <- reduceDimension(Mcds,
                        max_compongents = 2,
                        norm_method = 'log',
                        num_dim = 3,
                        reduction_method = 'tSNE',
                        verbose = F)

Mcds <- clusterCells(Mcds, verbose = F)

plot_cell_clusters(Mcds, color_by = 'Cluster')
plot_cell_clusters(Mcds, color_by = 'cluster_IDs')

Mcds_expresssed_genes <- row.names(subset(fData(Mcds),
                                          num_cells_expressed >=10))

# cluster_DEG <- differentialGeneTest(Mcds[Mcds_expresssed_genes,],
#                                     fullModelFormulaStr = '~seurat_clusters',
#                                     cores = 2)
# 
# Mcds_ordering_genes <- row.names(cluster_DEG)[order(cluster_DEG$qval)][1:1000]
# 
# Mcds <- setOrderingFilter(Mcds, ordering_genes = Mcds_ordering_genes)

Mcds <- reduceDimension(Mcds, method = "DDRTree")

Mcds <- orderCells(Mcds)

plot_cell_trajectory(Mcds, color_by = "cluster_IDs")

plot_cell_trajectory(Mcds, color_by = 'condition')
plot_cell_trajectory(Mcds, color_by = 'celltype')
```

```{r gran trajectory}

gcds <- cds[,cds$cluster_IDs %in% c('Stem Cell', paste0('Granulocyte',1:4))]
#dim(Mcds)

gcds <- detectGenes(gcds, min_expr = 0.1)

fData(gcds)$use_for_ordering <- fData(gcds)$num_cells_expressed > 0.05 * ncol(gcds)

gcds <- reduceDimension(gcds,
                        max_compongents = 2,
                        norm_method = 'log',
                        num_dim = 3,
                        reduction_method = 'tSNE',
                        verbose = F)

gcds <- clusterCells(gcds, verbose = F)

plot_cell_clusters(gcds, color_by = 'Cluster')
plot_cell_clusters(gcds, color_by = 'cluster_IDs')

gcds_expresssed_genes <- row.names(subset(fData(gcds),
                                          num_cells_expressed >=10))

# cluster_DEG <- differentialGeneTest(gcds[gcds_expresssed_genes,],
#                                     fullModelFormulaStr = '~seurat_clusters',
#                                     cores = 2)
# 
# gcds_ordering_genes <- row.names(cluster_DEG)[order(cluster_DEG$qval)][1:1000]
# 
# gcds <- setOrderingFilter(gcds, ordering_genes = gcds_ordering_genes)

gcds <- reduceDimension(gcds, method = "DDRTree")

gcds <- orderCells(gcds)

plot_cell_trajectory(gcds, color_by = "cluster_IDs")

plot_cell_trajectory(gcds, color_by = 'condition')
plot_cell_trajectory(gcds, color_by = 'celltype')

DimPlot(swbm, reduction = 'umap', label = T, repel = T) + NoLegend()

DimPlot(swbm, reduction = 'umap', label = T, repel = T, split.by = 'condition') + NoLegend()
```


```{r running with just MEPs}

MEPcds <- cds[,cds$cluster_IDs %in% c('MK')]

MEPcds <- estimateSizeFactors(MEPcds)
MEPcds <- estimateDispersions(MEPcds)

MEPcds <- detectGenes(MEPcds, min_expr = 0.1)

fData(MEPcds)$use_for_ordering <- fData(MEPcds)$num_cells_expressed > 0.05 * ncol(MEPcds)

MEPcds <- reduceDimension(MEPcds,
                        max_components = 2,
                        norm_method = 'log',
                        num_dim = 3,
                        reduction_method = 'tSNE',
                        verbose = F)

MEPcds <- clusterCells(MEPcds, verbose = F)

plot_cell_clusters(MEPcds, color_by = 'Cluster')
plot_cell_clusters(MEPcds, color_by = 'cluster_IDs')

MEPcds_expresssed_genes <- row.names(subset(fData(MEPcds),
                                          num_cells_expressed >=10))

# cluster_DEG <- differentialGeneTest(MEPcds[MEPcds_expresssed_genes,],
#                                     fullModelFormulaStr = '~seurat_clusters',
#                                     cores = 2)
# 
# MEPcds_ordering_genes <- row.names(cluster_DEG)[order(cluster_DEG$qval)][1:1000]
# 
# MEPcds <- setOrderingFilter(MEPcds, ordering_genes = MEPcds_ordering_genes)

MEPcds <- reduceDimension(MEPcds, method = "DDRTree")

MEPcds <- orderCells(MEPcds)

plot_cell_trajectory(MEPcds, color_by = "cluster_IDs")

plot_cell_trajectory(MEPcds, color_by = 'condition')
plot_cell_trajectory(MEPcds, color_by = 'celltype')

```

```{r adding seurat subclusters}

mks <- readRDS('data/MK-MEP_subclustering.rds')

# Checking to see if the cells are ordered the same way
summary(tstrsplit(pData(MEPcds)$cell, '-', keep = 1)[[1]] == rownames(mks@meta.data))

# Adding a column for the seurat subclusters
pData(MEPcds)[,'seurat_clusters'] <- mks$seurat_clusters

plot_cell_trajectory(MEPcds, color_by = 'seurat_clusters')
```

