---
title: "Monocle on Progenitors"
author: "D. Ford Hannum Jr."
date: "8/12/2020"
output: 
        html_document:
                toc: true
                toc_depth: 3
                number_sections: false
                theme: united
                highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(Seurat)
library(ggplot2)
library(data.table)
library(MAST)
library(monocle)
```

# Introduction

Looking at the monocle trajectory of the progenitors. Also looking at it when including MKs and ERYs. Following previous work done on monocle and following the [vignette](http://cole-trapnell-lab.github.io/monocle-release/docs/#constructing-single-cell-trajectories)

```{r loading data}
swbm <- readRDS('./data/wbm_clustered_filtered_named.rds')
```

```{r changing the levels of the data}
new_levels <- c('Granulocyte','Granulocyte','Granulocyte', 'B-cell','Progenitor', 
        'Granulocyte', 'Granulocyte','Monocyte','Macrophage','Erythroid','B-cell',
        'T-cell/NK', 'MK')
names(new_levels)  <- levels(swbm)
#new_levels
swbm <- RenameIdents(swbm, new_levels)

DimPlot(swbm, reduction = 'umap')
```

```{r creating monocle object}
mtx <- readMM('./data/filtered_feature_bc_matrix/matrix.mtx')
features <- read.table('./data/filtered_feature_bc_matrix/features.tsv')
colnames(features)[2] <- 'gene_short_name'

barcodes <- read.table('./data/filtered_feature_bc_matrix/barcodes.tsv')

pd <- new("AnnotatedDataFrame", data = barcodes)
fd <- new("AnnotatedDataFrame", data = features)

library(Matrix)

cds <- newCellDataSet(as(as.matrix(mtx),"sparseMatrix"),
                      phenoData = pd,
                      featureData = fd,
                      lowerDetectionLimit = 0.5,
                      expressionFamily = negbinomial.size())
```

```{r subsetting to valid cells} 

valid_cells <- colnames(swbm)

valid_cells <- paste0(valid_cells, '-1')

valid_cells <- rownames(subset(pData(cds),
                               V1 %in% valid_cells))

cds <- cds[,valid_cells]
```

```{r adding pData to cds object}
pData(cds)[,c('nCountRNA', 'nFeature_RNA', 'celltype','condition', 'percent.mt',
              'seurat_clusters', 'cluster_IDs')] <- swbm@meta.data[,c(2,3,5,6,8,11,15)]
colnames(pData(cds))[1] <- 'cell'
pData(cds)[,'identity'] <- Idents(swbm)
head(pData(cds),6)
```

# Focus on Progenitor Cluster

```{r prog cluster ordering}

prog <- cds[,cds$identity == 'Progenitor']

prog <- estimateSizeFactors(prog)
prog <- estimateDispersions(prog)

prog <- detectGenes(prog, min_expr = 0.1)

fData(prog)$use_for_ordering <- fData(prog)$num_cells_expressed > 0.05 * ncol(prog)

prog <- reduceDimension(prog,
                        max_compongents = 2,
                        norm_method = 'log',
                        num_dim = 3,
                        reduction_method = 'tSNE',
                        verbose = F)

prog <- clusterCells(prog, verbose = F)

plot_cell_clusters(prog, color_by = 'Cluster')

prog <- reduceDimension(prog, reduction_method = 'DDRTree')

prog <- reduceDimension()

prog <- orderCells(prog)

plot_cell_trajectory(prog)

plot_cell_trajectory(prog, color_by = 'Pseudotime')

plot_cell_trajectory(prog, color_by = 'State') +
        facet_wrap(~State, nrow = 3)

plot_cell_trajectory(prog, color_by = 'celltype')

plot_cell_trajectory(prog, color_by = 'Cluster')
```

## Checking of the Pseudotime Ordering

In the example they look at proliferation genes (Ccnb2, Myod1, Myog) and see how they're expressed over pseudotime. This also checks to see if the pseudotime is reversed or not.

These genes made sense for the example, but we should figure out our own (looking at the marker genes we've been using to distinguish progenitors).

### Example Genes

```{r checking proliferation markers}
blast_genes <- row.names(subset(fData(prog), 
                                gene_short_name %in% c('Mpo','Ank1','Vwf')))
plot_genes_jitter(prog[blast_genes,],
                  grouping = 'State',
                  min_expr = 0.1)
plot_genes_violin(prog[blast_genes,],
                  grouping = 'State',
                  min_expr = 0.1)
```

They also checked markers of myogenic progress.

```{r myogenic progress}
prog_expressed_genes <- row.names(subset(fData(prog), num_cells_expressed >=10))

prog_filtered <- prog[prog_expressed_genes,]

my_genes <- row.names(subset(fData(prog_filtered),
                              gene_short_name %in% c('Itga2b','Gata1','Mpo', 'Dhrs3',
                                                     'Vwf','Myb','Cd44','Kit')))

prog_subset <- prog_filtered[my_genes,]


featureNames(prog_subset) <- fData(prog_subset)$gene_short_name
sampleNames(prog_subset) <- prog_subset$cell

plot_genes_in_pseudotime(prog_subset[1:3,])
plot_genes_in_pseudotime(prog_subset[4:6,])
plot_genes_in_pseudotime(prog_subset[7:8,])
```

# DE By Trajectory States

```{r de for trajectory states}

diff_test_res <- differentialGeneTest(prog,
                                      fullModelFormulaStr = '~State')

diff_test_res <- diff_test_res[diff_test_res$qval < 0.05,]
head(diff_test_res)

```

