---
title: "v2.5.2.MEP.Introspection.part2"
author: "D. Ford Hannum Jr."
date: "11/18/2020"
output: 
        html_document:
                toc: true
                toc_depth: 3
                number_sections: false
                theme: united
                highlight: tango
---

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(Seurat)
library(ggplot2)
library(data.table)
library(MAST)
library(SingleR)
library(dplyr)
library(tidyr)
library(limma)
library(scRNAseq)
```

# Data Intro and Initial Processing

```{r loading data}
wbm <- readRDS('./data/v2/lesser.combined.integrated.NAMED.rds')
```

```{r changing labels}
# Changing HSPC to Multipotent Progenitor

# Changing CMP to Progenitors?

wbm$idents2 <- ifelse(wbm$idents != 'HSPC' & wbm$idents != 'CMP' & wbm$idents != 'MEP/Mast', as.character(wbm$idents),
                      ifelse(wbm$idents == 'MEP/Mast', 'MEP/MCP',
                             ifelse(wbm$idents == 'HSPC','MPP', 'Progenitor')))

Idents(wbm) <- wbm$idents2

DefaultAssay(wbm) <- 'RNA'
```

```{r basic UMAP plot}
color_pal <- c("#0072B2", "#CC79A7", "#009E73", "#56B4E9","#D55E00",
               "#E69F00","#999999", 'violet',"red", 'black')

DimPlot(wbm,reduction = 'umap', label = T, repel = T, cols = color_pal)
```

# Introduction

Things that were brought up last meeting:

**What are the MEP/Mast cells, are these really MEPs?**

1. Look at Krause markers?
2. Look at markers for MEPs?
3. Coexpression of different markers
4. How can we isolate these cells?

# Krause Markers

```{r mep gene lists}
mep.genes <- c('Klf1','Fli1','Gata1','Gata2','Runx1','Tal1','Myb')

mep.genes2 <- c('Hpgds','Smox','Lbh','Nr4a1','Gpr141')

cmp.genes <- c('Spink2','Csf3r','Ighm','Glipr1','Miat')

erp.genes <- c('Rhag','Ahspp','Klf1')

mkp.genes <- c('Pf4','Vwf','Trpc6','Gp6','Itgb3','Fcer1g','Selp')

mep.erp.genes <- c('Cnrip1','Ank1','Tfr2','Acsm3','Cpa3')

mep.mkp.genes <- c('C3orf58','Pdia5','Rgs18','Pdgfa','Plek','Arhgap6')

cmp.mep.genes <- c('Cd52','Id3','Smim24')

mep.mkp.erp.genes <- c('Abcc4','C6orf25','Casp3','Ibtk','Gata1')
```

```{r dot plots}
DotPlot(wbm, features = mep.genes) + coord_flip() + ggtitle ('MEP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = mep.genes2) + coord_flip() + ggtitle ('MEP Genes #2') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = cmp.genes) + coord_flip() + ggtitle ('CMP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = erp.genes) + coord_flip() + ggtitle ('ERP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = mkp.genes) + coord_flip() + ggtitle ('MKP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = mep.erp.genes) + coord_flip() + ggtitle ('MEP/ERP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = mep.mkp.genes) + coord_flip() + ggtitle ('MEP/MKP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = cmp.mep.genes) + coord_flip() + ggtitle ('CMP/MEP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))

DotPlot(wbm, features = mep.mkp.erp.genes) + coord_flip() + ggtitle ('MEP/MKP/ERP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Markers of Interest

```{r key ones from above}
moi <- c('Casp3','Cpa3','Runx1')

DotPlot(wbm, features = moi) + coord_flip() + ggtitle ('Strongest Markers') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


# MEP Markers

Other markers that I have from priya and other sources.

```{r otros}
otro.mep.markers <- c('Cd34','Cd38')

DotPlot(wbm, features = otro.mep.markers) + coord_flip() + # ggtitle ('MEP Genes') +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Neither of these was useful

# Coexpression of different markers
 
Looking at some of the top hits (Mcpt8, Prss34) and other MEP genes (Gata2, Runx1, Cpa3), to see how they overlap

```{r data for goi}
goi <- c('Itga2b','Mcpt8','Prss34','Gata2','Runx1','Cpa3')
rnas <- as.data.frame(t(wbm@assays$RNA[goi,]))
rnas$cluster <- wbm$idents2
mep.mcp <- rnas[rnas$cluster == 'MEP/MCP',]
```


```{r coexpression data}
combos <- combn(goi,2)

for (i in 1:ncol(combos)){
        x <- combos[1,i]
        y <- combos[2,i]
        # print(paste(x,y))
        
        temp <- mep.mcp[, colnames(mep.mcp) %in% c(x,y)]
        
        colnames(temp) <- c('V1','V2')
        
        # temp2 <- lm(V1 ~ V2, temp)
        # r2 <- round(summary(temp2)$r.squared,2)
        
        print(ggplot(temp, aes(x = V1, y = V2)) + geom_point() + 
                             theme_bw() + xlab(x) + ylab(y))
        
        # print(paste0('The r^2 value is : ',r2))
        
        tbl <- table(mep.mcp[,y] >0, mep.mcp[,x] > 0)
        
        colnames(tbl) <- c(paste0(x, ' = 0'), paste0(x, ' > 0'))
        rownames(tbl) <- c(paste0(y, ' = 0'), paste0(y, ' > 0'))
        print(tbl)
}
```

# How to select this MEP/MCP population

## First Step

Most of these cells are coming from the enriched Mpl experiment so I would start there.

## Second Step

```{r defining by markers here first}
# rnas$exp <- wbm$State
# w <- rnas[rnas$exp == 'enrMpl',]
# temp <- table(w$Itga2b > 0, w$cluster)
# rownames(temp) <- c("No Itga2b", "Itga2b Expresion")
# temp
enrMpl <- subset(wbm, State == 'enrMpl')

mrks <- FindMarkers(enrMpl,
                    ident.1 = 'MEP/MCP',
                    min.pct = .9,
                    min.diff.pct = .5)

head(mrks)
```

```{r looking at these genes in violins}
mrks$diff <- mrks$pct.1 - mrks$pct.2

mrks <- mrks[order(mrks$diff, decreasing = T),]

for (i in rownames(head(mrks))){
        print(VlnPlot(enrMpl, features = i))
}

nonEnrMpl <- subset(wbm, enriched == FALSE)
nbeal <- subset(wbm, State == 'Nbeal_cntrl')
for (i in rownames(head(mrks))){
        print(VlnPlot(nbeal, features = i))
}
```

