---
title: "v3.WriteUp.Subsets"
author: "D. Ford Hannum Jr."
date: "12/17/2020"
output: 
        html_document:
                toc: true
                toc_depth: 3
                number_sections: false
                theme: united
                highlight: tango
---

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(Seurat)
library(ggplot2)
library(data.table)
library(MAST)
library(SingleR)
library(dplyr)
library(tidyr)
library(limma)
library(scRNAseq)
library(ggpubr)
library(xlsx)
```


```{r}
wbm <- readRDS('./data/v3/mpl.migr1.121420-2.rds')

color_pal <- c("#0072B2", "#CC79A7", "#009E73", "#56B4E9","#D55E00",
               "#E69F00","#999999", 'violet',"red", 'black')

DimPlot(wbm, label = T, repel = T, group.by = 'v7.labels',
        cols = color_pal)
```

# Further Analysis Subsetting

For further analysis we are going to exclude the Granulocyte and Lympthocyte Clusters, becuase they area driving the clustering and UMAP projection.

Would it even make a difference??????

```{r}

wo.GranBcell <- subset(wbm, v7.labels %in% c('MEP/MCP','Erythroid','MEP/ERP','MK'))

wo.GranBcell <- NormalizeData(wo.GranBcell, normalization.method = 'LogNormalize', scale.factor = 10000)

wo.GranBcell <- FindVariableFeatures(wo.GranBcell, selection.method = 'vst', nfeatures = 2000)

wo.GranBcell <- ScaleData(wo.GranBcell, features = rownames(wo.GranBcell))

wo.GranBcell <- RunPCA(wo.GranBcell, features = VariableFeatures(wo.GranBcell), verbose =  F)

ElbowPlot(wo.GranBcell)

DimHeatmap(wo.GranBcell, dims = 1, cells = 100, balanced = T)

DimPlot(wo.GranBcell, reduction = 'pca', dims = 1:2)

# Going with 8

wo.GranBcell <- FindNeighbors(wo.GranBcell, dims = 1:10)

for (i in seq(0,1,by = .1)){
  wo.GranBcell <- FindClusters(wo.GranBcell, resolution = i, verbose = F)
}

wo.GranBcell <- RunUMAP(wo.GranBcell, dims = 1:10)

wo.GranBcell <- RunTSNE(wo.GranBcell, dims = 1:10)

rna.col.names <- colnames(wo.GranBcell@meta.data)[grepl('RNA_snn', 
                                                        colnames(wo.GranBcell@meta.data), ignore.case = T)]
# rna.col.names

# for (i in rna.col.names){
#   print(DimPlot(wo.GranBcell, group.by = i, label = T, repel = T) + ggtitle(i))
# }

Idents(wo.GranBcell) <- wo.GranBcell$RNA_snn_res.0.5

DimPlot(wo.GranBcell, repel = T, label = T)

DimPlot(wo.GranBcell, repel = T, label = T, split.by = 'v7.labels', ncol = 2)
```

```{r}

temp.subset <- subset(wbm, v7.labels %in% c('MEP/MCP','Erythroid','MEP/ERP','MK'))

DimPlot(temp.subset, repel = T, label = T) + ggtitle('How these cells looked in the original projection')

DimPlot(wo.GranBcell, label = T, repel = T, group.by = 'v6.labels')
```


## Cross-Tabulation

```{r}
old.clusters <- Idents(temp.subset)
new.clusters <- Idents(wo.GranBcell)


ct <- data.frame(old.clusters = old.clusters,
                 new.clusters = new.clusters)

temp.df <- as.data.frame(table(ct$old.clusters, ct$new.clusters))

temp.df$Var1.total <- NA
for (i in unique(temp.df$Var1)){
  temp.df[temp.df$Var1 == i,]$Var1.total <- sum(temp.df[temp.df$Var1 == i,]$Freq)
}
temp.df$Var1.perc <- temp.df$Freq /temp.df$Var1.total

temp.df$Var2.total <- NA
for (i in unique(temp.df$Var2)){
  temp.df[temp.df$Var2 == i,]$Var2.total <- sum(temp.df[temp.df$Var2 == i,]$Freq)
}
temp.df$Var2.perc <- temp.df$Freq /temp.df$Var2.total

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var1.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Original Clusters') +
  ylab('New Clusters') +
  ggtitle('Filled by Column Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var2.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Original Clusters') +
  ylab('New Clusters') +
  ggtitle('Filled by Row Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')
```

Seems to be a lot of overlap with Eryth and MK cells

```{r}
FeaturePlot(wo.GranBcell, features = c('Vwf','Itga2b'))
```

Looks like MKs are part of that branch in the cluster shared with Erythrocytes, may want to trim that one off :)

# Looking at the Flowy Markers Priya Used

```{r}
FeaturePlot(wo.GranBcell, features = 'Kit')

# matrix(wo.GranBcell@assays$RNA['Kit',])[,1] > 0

wo.GranBcell$kit.expression <- ifelse(matrix(wo.GranBcell@assays$RNA['Kit',])[,1] > 0,
                                      1,0)
wo.GranBcell$cd34.expression <- ifelse(matrix(wo.GranBcell@assays$RNA['Cd34',])[,1] > 0,
                                      1,0)
wo.GranBcell$itga2b.expression <- ifelse(matrix(wo.GranBcell@assays$RNA['Itga2b',])[,1] > 0,
                                      1,0)

wo.GranBcell$cd16.expression <- ifelse(matrix(wo.GranBcell@assays$RNA['Fcgr3',])[,1] > 0,
                                      1,0)

wo.GranBcell$cd32.expression <- ifelse(matrix(wo.GranBcell@assays$RNA['Fcgr2b',])[,1] > 0,
                                      1,0)
```


```{r}
wo <- wo.GranBcell

FeaturePlot(wo, features = 'kit.expression')

VlnPlot(wo, features = 'Kit')

FeaturePlot(wo, features = c("Fcgr3"))

FeaturePlot(wo, features = c("Fcgr2b"))

FeaturePlot(wo, features = 'Itga2b')

FeaturePlot(wo, features = 'Cd34')
# FeaturePlot(wbm, features = 'Cd34')

FeaturePlot(wo, features = 'Ly6a')
```

## Using Cd41, Cd16, and Cd32

```{r}

wo$cd41.pos.mep <- ifelse(wo$itga2b.expression == 1 & wo$cd16.expression == 0 & wo$cd32.expression == 0, 1,0)

wo$cd41.neg.mep <- ifelse(wo$itga2b.expression == 0 & wo$cd16.expression == 0 & wo$cd32.expression == 0, 1,0)

wo$mkp <- ifelse(wo$itga2b.expression == 1 & wo$cd16.expression == 1 & wo$cd32.expression == 1, 1,0)

wo$mep.mkp.type <- ifelse(wo$cd41.pos.mep == 1, 'Cd41.pos.MEP',
                          ifelse(wo$cd41.neg.mep == 1, 'Cd41.neg.MEP',
                                 ifelse(wo$mkp == 1, 'MKP','Other')))

summary(as.factor(wo$mep.mkp.type))

DimPlot(wo, split.by = 'mep.mkp.type', ncol = 2, label = T, repel = T)


```

### CrossTab

```{r}
table(wo$mep.mkp.type, wo$seurat_clusters)

temp.df <- as.data.frame(table(wo$mep.mkp.type, wo$seurat_clusters))

temp.df$Var1.total <- NA
for (i in unique(temp.df$Var1)){
  temp.df[temp.df$Var1 == i,]$Var1.total <- sum(temp.df[temp.df$Var1 == i,]$Freq)
}
temp.df$Var1.perc <- temp.df$Freq /temp.df$Var1.total

temp.df$Var2.total <- NA
for (i in unique(temp.df$Var2)){
  temp.df[temp.df$Var2 == i,]$Var2.total <- sum(temp.df[temp.df$Var2 == i,]$Freq)
}
temp.df$Var2.perc <- temp.df$Freq /temp.df$Var2.total

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var1.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('New Clusters') +
  ggtitle('Filled by Column Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var2.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('New Clusters') +
  ggtitle('Filled by Row Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')
```

### Comparing from Mpl to Migr1

```{r}
wo$Experiment <- ifelse(wo$Condition2 %in% c('Migr1','enrMigr1'), 'Migr1','Mpl')
table(wo$mep.mkp.type, wo$Experiment)

temp.df <- as.data.frame(table(wo$mep.mkp.type, wo$Experiment))

temp.df$Var1.total <- NA
for (i in unique(temp.df$Var1)){
  temp.df[temp.df$Var1 == i,]$Var1.total <- sum(temp.df[temp.df$Var1 == i,]$Freq)
}
temp.df$Var1.perc <- temp.df$Freq /temp.df$Var1.total

temp.df$Var2.total <- NA
for (i in unique(temp.df$Var2)){
  temp.df[temp.df$Var2 == i,]$Var2.total <- sum(temp.df[temp.df$Var2 == i,]$Freq)
}
temp.df$Var2.perc <- temp.df$Freq /temp.df$Var2.total

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var1.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('Experiment') +
  ggtitle('Filled by Column Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var2.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('Experimen') +
  ggtitle('Filled by Row Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')
```

This is still really raw with many cells being labeled as others. Perhaps if we identify clusters then we would have a better idea about labels and then we look at this same figure again, but instead with the named clusters.


## Using Cd41, Cd16, Cd32 and Cd34

Adding the marker of Cd34^+^ for MkPs. This may not work great because we are seeing very few cells expressing Cd34 in the global data.

```{r}
FeaturePlot(wbm, features = 'Cd34')

summary(matrix(wbm@assays$RNA['Cd34',])[,1] > 0)
summary(as.factor(wo$cd34.expression))
```

We are only seeing 86 cells in the entire experiment with any expression of Cd34, and 23 of these are within this subsample. 

```{r}

wo$cd41.pos.mep <- ifelse(wo$itga2b.expression == 1 & wo$cd16.expression == 0 & 
                            wo$cd32.expression == 0 & wo$cd34.expression == 0, 1,0)

wo$cd41.neg.mep <- ifelse(wo$itga2b.expression == 0 & wo$cd16.expression == 0 & 
                            wo$cd32.expression == 0 & wo$cd34.expression == 0, 1,0)

wo$mkp <- ifelse(wo$itga2b.expression == 1 & wo$cd16.expression == 1 & 
                   wo$cd32.expression == 1 & wo$cd34.expression == 1, 1,0)

wo$mep.mkp.type <- ifelse(wo$cd41.pos.mep == 1, 'Cd41.pos.MEP',
                          ifelse(wo$cd41.neg.mep == 1, 'Cd41.neg.MEP',
                                 ifelse(wo$mkp == 1, 'MKP','Other')))

summary(as.factor(wo$mep.mkp.type))

DimPlot(wo, split.by = 'mep.mkp.type', ncol = 2, label = T, repel = T)


```

### CrossTab

```{r}
table(wo$mep.mkp.type, wo$seurat_clusters)

temp.df <- as.data.frame(table(wo$mep.mkp.type, wo$seurat_clusters))

temp.df$Var1.total <- NA
for (i in unique(temp.df$Var1)){
  temp.df[temp.df$Var1 == i,]$Var1.total <- sum(temp.df[temp.df$Var1 == i,]$Freq)
}
temp.df$Var1.perc <- temp.df$Freq /temp.df$Var1.total

temp.df$Var2.total <- NA
for (i in unique(temp.df$Var2)){
  temp.df[temp.df$Var2 == i,]$Var2.total <- sum(temp.df[temp.df$Var2 == i,]$Freq)
}
temp.df$Var2.perc <- temp.df$Freq /temp.df$Var2.total

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var1.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('New Clusters') +
  ggtitle('Filled by Column Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var2.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('New Clusters') +
  ggtitle('Filled by Row Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')
```

### Comparing from Mpl to Migr1

```{r}
wo$Experiment <- ifelse(wo$Condition2 %in% c('Migr1','enrMigr1'), 'Migr1','Mpl')
table(wo$mep.mkp.type, wo$Experiment)

temp.df <- as.data.frame(table(wo$mep.mkp.type, wo$Experiment))

temp.df$Var1.total <- NA
for (i in unique(temp.df$Var1)){
  temp.df[temp.df$Var1 == i,]$Var1.total <- sum(temp.df[temp.df$Var1 == i,]$Freq)
}
temp.df$Var1.perc <- temp.df$Freq /temp.df$Var1.total

temp.df$Var2.total <- NA
for (i in unique(temp.df$Var2)){
  temp.df[temp.df$Var2 == i,]$Var2.total <- sum(temp.df[temp.df$Var2 == i,]$Freq)
}
temp.df$Var2.perc <- temp.df$Freq /temp.df$Var2.total

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var1.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('Experiment') +
  ggtitle('Filled by Column Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')

ggplot(temp.df, aes(x = Var1, y = Var2, fill = Var2.perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Surface Marker ID') +
  ylab('Experimen') +
  ggtitle('Filled by Row Percentages') + theme_bw() +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')
```



# Cell Cycle Work

```{r}
ccg <- read.table('/Users/dfhannum/Desktop/CZI_Ovary/data/CellCycleGeneList_1134.txt', header = T, row.names = 1, sep = '\t')

id <- read.table('/Users/dfhannum/Desktop/CZI_Ovary/data/genesymbol.all.list', header = F, sep = '\t')

ccg$gene_name <- tstrsplit(ccg$NAME, ' ', keep =1)[[1]]

# summary(id$V1 %in% ccg$gene_name)

ccg$V2 <- rownames(ccg)

data <- merge(ccg, id, by = 'V2')

data$gene <- NA

wo.genes <- toupper(rownames(wo))

data$gene <- ifelse(data$V1 %in% wo.genes, data$V1,
                    ifelse(data$gene_name %in% wo.genes,
                           data$gene_name, NA))

data <- data[!is.na(data$gene),]

cnts <- wo@assays$RNA

len <- length(colnames(wo))

x <- data.frame(row.names = colnames(wo),
           cell.count.sum = numeric(len),
           g1.s.raw.score = numeric(len),
           g2.raw.score = numeric(len),
           g2.m.raw.score = numeric(len),
           s.raw.score = numeric(len),
           m.g1.raw.score = numeric(len),
           g1.s.norm.score = numeric(len),
           g2.norm.score = numeric(len),
           g2.m.norm.score = numeric(len),
           s.norm.score = numeric(len),
           m.g1.norm.score = numeric(len)
           )

cnt <- 2


data$gene2 <- NA
for (i in 1:dim(data)[1]){
        temp.gene <- data$gene[i]
        temp.row <- rownames(cnts)[grep(temp.gene, rownames(cnts), ignore.case = T)]
        data$gene2[i] <- temp.row[1]
        # print(paste(temp.gene, temp.row))
}

for (i in unique(data$PHASE)[1:5]){
  # print(i)
  # print(colnames(x)[cnt])
  
  phase.genes <- data[data$PHASE == i,]$gene2
  # print(phase.genes)
  temp <- cnts[rownames(cnts) %in% phase.genes,]
  temp.sums <- colSums(temp)
  x[,cnt] <- temp.sums
  cnt <- cnt + 1
}

x$cell.count.sum <- colSums(cnts)
head(x)

for (i in 2:6){
  # print(colnames(x)[i])
  # print(colnames(x)[i+5])
  
  temp_mean <- mean(x[,i])
  temp_sd <- sd(x[,i])
  x[,i+5] <- (x[,i] - temp_mean) / temp_sd
}

x$max_score <- colnames(x[,7:11])[apply(x[,7:11],1,which.max)]
```

```{r}
x$idents <- wo$seurat_clusters

x$max <- ifelse(x$max_score == 'g1.s.norm.score', 'G1/S',
                ifelse(x$max_score == 'g2.m.norm.score', 'G2/M',
                       ifelse(x$max_score == 'g2.norm.score', 'G2',
                              ifelse(x$max_score == 'm.g1.norm.score', 'M/G1', 'S'))))

wo$cell.cycle.phase <- x$max

df <- as.data.frame(table(x$max, x$idents))
df$cluster_count <- NA

for(i in df$Var2){
  df[df$Var2 == i,]$cluster_count <- sum(df[df$Var2 == i, 'Freq'])
}

df$perc <- df$Freq / df$cluster_count

ggplot(df, aes (x = Var1, y = Var2, fill = perc)) + geom_tile() +
  geom_text(aes(label = Freq)) + xlab('Cell Cycle') + ylab ('Cluster') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .2)
```

# Fibrotic Scores

```{r}
wo <- readRDS('./data/v3/woGranBcell.rds')
```


```{r}
fib.genes <- read.table('./data/profibrotic_factors.txt')
fib.genes <- fib.genes$V1[-1]

summary(fib.genes %in% rownames(wo))
```

35 of the 43 fibrotic genes are contained in our data.

```{r}
cnts <- wo@assays$RNA

cell.total.score <- colSums(cnts)

cnts <- cnts[rownames(cnts) %in% fib.genes,]

cell.raw.score <- colSums(cnts)

fib.df <- data.frame(cell = colnames(cnts),
                     cell.total.score = cell.total.score,
                     fib.total.score = cell.raw.score)

fib.df$weighted.score <- fib.df$fib.total.score / fib.df$cell.total.score

fib.df$scaled.score <- (fib.df$weighted.score - mean(fib.df$weighted.score)) /
  sd(fib.df$weighted.score)

wo$fibrosis.score <- fib.df$scaled.score

FeaturePlot(wo, features = 'fibrosis.score')
```



```{r}
# saveRDS(wo.GranBcell,'./data/v3/woGranBcell.rds')
```


# Looking at interesting islands

```{r}
wo <- readRDS('./data/v3/woGranBcell.rds')

DimPlot(wo, label = T, repel = T)
```

```{r}
nine.markers <- FindMarkers(wo,
                             ident.1 = 9,
                             test.use = 'MAST',
                             logfc.threshold = log(2))
eight.markers <- FindMarkers(wo,
                             ident.1 = 8,
                             test.use = 'MAST',
                             logfc.threshold = log(2))


nine.markers <- nine.markers[nine.markers$p_val_adj < 0.05,]
eight.markers <- eight.markers[eight.markers$p_val_adj < 0.05,]

nine.markers <- nine.markers[order(nine.markers$avg_logFC, decreasing = T),]
eight.markers <- eight.markers[order(eight.markers$avg_logFC, decreasing = T),]

head(nine.markers)

```

*Cma1*: mast cell marker
*Tpsb2*: mast cell marker
*Cpa3*: mast cell gene
*Gzmb*: expressed by t-cell and NK cells
*Kit*: proliferation marker

```{r}
head(eight.markers)
```

*Igkc*: produced by b-cells
*Ighm*: b-cells
*Ly6d*: may be a specification marker between b-cells and t-cell development. marks the earliest stage of b-cell specification.

## Full DE

```{r}
de.genes <- data.frame(p_val = numeric(),
                       avg_logFC = numeric(),
                       pct.1 = numeric(),
                       pct.2 = numeric(),
                       p_val_adj = numeric(),
                       cluster = character(),
                       gene = character())

for (i in 0:9){
  print(i)
  x <- FindMarkers(wo,
                   ident.1 = i,
                   test.use = 'MAST',
                   logfc.threshold = log(2),
                   only.pos = F)
  x$cluster <- i
  x$gene <- rownames(x)
  x <- x[x$p_val_adj < 0.05,]
  x <- x[abs(x$avg_logFC) > log(2),]
  de.genes <- rbind(de.genes,x)
}

summary(as.factor(de.genes$cluster))
# table(rownames(de.genes))

out <- split(de.genes, f = de.genes$cluster)

library(xlsx)

for (i in 1:length(out)){
  print(i)
  print(unique(out[[i]]$cluster))
}



for (i in 1:length(out)){
  print(i)
  temp.name <- unique(out[[i]]$cluster)
  temp.name <- gsub('/','_',temp.name)
  print(temp.name)
  write.xlsx(x = out[[i]], file = "data/v3/MEP.subset.between.cluster.markers.xlsx",
             sheetName=temp.name, append = T)
}
write.xlsx(x = de.genes, file = "data/v3/MEP.subset.between.cluster.markers.xlsx",
           sheetName='All-Clusters', append = T)


```



# Figures

```{r}
wo <-readRDS('./data/v3/woGranBcell.rds')

wo$Experiment <- ifelse(wo$Condition2 %in% c('Migr1','enrMigr1'),
                        'MigR1','MPLW515L')

wo$figure.labels <- ifelse(wo$v7.labels == 'MEP/ERP', 'MEP-2',
                           ifelse(wo$v7.labels == 'MEP/MCP', 'MEP-1',
                                  ifelse(wo$v7.labels == 'MK','MK', "Ery")))

```

## a-c) UMAPs



```{r}
DimPlot(wo, label = T, repel = T, pt.size = 0.01) +
  theme_bw() +
  theme(text = element_text(size = 6, family = 'sans'))

ggsave('./figures/version7.for.v3/2a.umap.pdf', device = 'pdf',
       height = 2.5, width = 3, units = 'in', dpi = 600)
```

```{r}


DimPlot(wo, pt.size = 0.01, split.by = 'Experiment') +
  theme_bw() + NoLegend() +
  theme(text = element_text(size = 6, family = 'sans'))

ggsave('./figures/version7.for.v3/2b.umap.split.by.exp.pdf', device = 'pdf',
       height = 2.5, width = 4, units = 'in', dpi = 600)
```

```{r}
DimPlot(wo, pt.size = 0.01, split.by = 'figure.labels') +
  theme_bw() + NoLegend() +
  theme(text = element_text(size = 6, family = 'sans'))

ggsave('./figures/version7.for.v3/2c.umap.split.by.labels.pdf', device = 'pdf',
       height = 2.5, width = 7, units = 'in', dpi = 600)
```

## d) MK subclusters full UMAP

```{r}
wbm$mk.ident <- ifelse(wbm$figure.labels == 'MK',1,0)

FeaturePlot(wbm, features = 'mk.ident')

DimPlot(wbm, label = T, repel = T,
        group.by = 'figure.labels', 
        cols = c(rep('grey',7),'red','grey','grey'), pt.size = 0.01) +
  theme_bw() + ggtitle('') + NoLegend() +
  theme(text = element_text(size = 6, family = 'sans'))

ggsave('./figures/version7.for.v3/2d.full.UMAP.highlight.MKs.pdf', device = 'pdf',
       height = 2.5, width = 3, units = 'in', dpi = 600)
```


# Comparing two MK populations

```{r}
temp.subset <- subset(wo, figure.labels == 'MK')


mk.different.markers <- FindMarkers(temp.subset,
            ident.1 = 5,
            ident.2 = 1,
            test.use = 'MAST',
            logfc.threshold = log(2))

mk.different.markers <- mk.different.markers[mk.different.markers$p_val_adj < 0.05,]

mk.different.markers <- mk.different.markers[order(mk.different.markers$avg_logFC, decreasing = T),]
dim(mk.different.markers)
head(mk.different.markers)
```

```{r}
tail(mk.different.markers)
```

```{r}
DimPlot(temp.subset, label = T, repel = T)
```

```{r}
table(Idents(temp.subset), temp.subset$Condition2)
VlnPlot(temp.subset, features = c('Itga2b','Pf4','Vwf', 'Gp6'), 
        ncol = 2, pt.size = 0)
VlnPlot(temp.subset, features = c('Gp9'), pt.size = 0)
```
```{r}
temp.subset$var2 <- Idents(temp.subset)

temp.subset2 <- subset(temp.subset, var2 == 5)

dim(temp.subset2)

subcluster.5.MK.cells <- colnames(temp.subset2)

subcluster.5.MK.cells

DimPlot(wbm, cells = subcluster.5.MK.cells)

wbm$subcluster.5.MK.cells <- ifelse(colnames(wbm) %in% subcluster.5.MK.cells, 1, 0)

FeaturePlot(wbm, features = 'subcluster.5.MK.cells')
```


## Are they doublets?

```{r}
VlnPlot(wo, features = 'nFeature_RNA')
VlnPlot(wo, features = 'nCount_RNA')

VlnPlot(wo, features = 'nFeature_RNA', split.by = 'figure.labels')
VlnPlot(wo, features = 'nCount_RNA', split.by = 'figure.labels')
```

