---
title: "v3.WriteUp.Extras"
author: "D. Ford Hannum Jr."
date: "12/17/2020"
output: 
        html_document:
                toc: true
                toc_depth: 3
                number_sections: false
                theme: united
                highlight: tango
---

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(Seurat)
library(ggplot2)
library(data.table)
library(MAST)
library(SingleR)
library(dplyr)
library(tidyr)
library(limma)
library(scRNAseq)
library(ggpubr)
library(xlsx)
```


```{r}
wbm <- readRDS('./data/v3/mpl.migr1.121420-2.rds')
```

To Do:

* Cell Cycle Analysis
* Profibrotic Score
* Compare to ET and PV to PMF DE list
* Trajectory

# Cell Cycle Analysis

```{r}
ccg <- read.table('/Users/dfhannum/Desktop/CZI_Ovary/data/CellCycleGeneList_1134.txt', header = T, row.names = 1, sep = '\t')

id <- read.table('/Users/dfhannum/Desktop/CZI_Ovary/data/genesymbol.all.list', header = F, sep = '\t')

ccg$gene_name <- tstrsplit(ccg$NAME, ' ', keep =1)[[1]]

# summary(id$V1 %in% ccg$gene_name)

ccg$V2 <- rownames(ccg)

data <- merge(ccg, id, by = 'V2')

data$gene <- NA

wbm.genes <- toupper(rownames(wbm))

data$gene <- ifelse(data$V1 %in% wbm.genes, data$V1,
                    ifelse(data$gene_name %in% wbm.genes,
                           data$gene_name, NA))

data <- data[!is.na(data$gene),]

cnts <- wbm@assays$RNA

len <- length(colnames(wbm))

x <- data.frame(row.names = colnames(wbm),
           cell.count.sum = numeric(len),
           g1.s.raw.score = numeric(len),
           g2.raw.score = numeric(len),
           g2.m.raw.score = numeric(len),
           s.raw.score = numeric(len),
           m.g1.raw.score = numeric(len),
           g1.s.norm.score = numeric(len),
           g2.norm.score = numeric(len),
           g2.m.norm.score = numeric(len),
           s.norm.score = numeric(len),
           m.g1.norm.score = numeric(len)
           )

cnt <- 2


data$gene2 <- NA
for (i in 1:dim(data)[1]){
        temp.gene <- data$gene[i]
        temp.row <- rownames(cnts)[grep(temp.gene, rownames(cnts), ignore.case = T)]
        data$gene2[i] <- temp.row[1]
        # print(paste(temp.gene, temp.row))
}

for (i in unique(data$PHASE)[1:5]){
  # print(i)
  # print(colnames(x)[cnt])
  
  phase.genes <- data[data$PHASE == i,]$gene2
  # print(phase.genes)
  temp <- cnts[rownames(cnts) %in% phase.genes,]
  temp.sums <- colSums(temp)
  x[,cnt] <- temp.sums
  cnt <- cnt + 1
}

x$cell.count.sum <- colSums(cnts)
head(x)

for (i in 2:6){
  # print(colnames(x)[i])
  # print(colnames(x)[i+5])
  
  temp_mean <- mean(x[,i])
  temp_sd <- sd(x[,i])
  x[,i+5] <- (x[,i] - temp_mean) / temp_sd
}

x$max_score <- colnames(x[,7:11])[apply(x[,7:11],1,which.max)]
```

```{r}
x$idents <- wbm$v6.labels

x$max <- ifelse(x$max_score == 'g1.s.norm.score', 'G1/S',
                ifelse(x$max_score == 'g2.m.norm.score', 'G2/M',
                       ifelse(x$max_score == 'g2.norm.score', 'G2',
                              ifelse(x$max_score == 'm.g1.norm.score', 'M/G1', 'S'))))

wbm$cell.cycle.phase <- x$max

df <- as.data.frame(table(x$max, x$idents))
df$cluster_count <- NA

for(i in df$Var2){
  df[df$Var2 == i,]$cluster_count <- sum(df[df$Var2 == i, 'Freq'])
}

df$perc <- df$Freq / df$cluster_count

ggplot(df, aes (x = Var1, y = Var2, fill = perc)) + geom_tile() +
  theme_bw() +
  geom_text(aes(label = Freq)) + xlab('Cell Cycle') + ylab ('Cluster') +
  scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .2)
```

```{r}
DimPlot(wbm, group.by = 'cell.cycle.phase')
DimPlot(wbm, group.by = 'v6.labels', label = T)
```

# Profibrotic score

## First Gene Set

A list of 43 genes provided by Priya

```{r}
fib.genes <- read.table('./data/profibrotic_factors.txt')
fib.genes <- fib.genes$V1[-1]

summary(fib.genes %in% rownames(wbm))
```

35 of the 43 fibrotic genes are contained in our data.

```{r}
cnts <- wbm@assays$RNA

cell.total.score <- colSums(cnts)

cnts <- cnts[rownames(cnts) %in% fib.genes,]

cell.raw.score <- colSums(cnts)

fib.df <- data.frame(cell = colnames(cnts),
                     cell.total.score = cell.total.score,
                     fib.total.score = cell.raw.score)

fib.df$weighted.score <- fib.df$fib.total.score / fib.df$cell.total.score

ggplot(fib.df, aes (x = weighted.score)) + geom_density()

fib.df$scaled.score <- (fib.df$weighted.score - mean(fib.df$weighted.score)) /
  sd(fib.df$weighted.score)

wbm$fibrosis.score <- fib.df$scaled.score
wbm$fibrosis.score.unstandardized <- fib.df$weighted.score

FeaturePlot(wbm, features = 'fibrosis.score', cols = c('lightgrey','red'))
FeaturePlot(wbm, features = 'fibrosis.score.unstandardized')
```

### Looking at each large/generic cluster as a whole

```{r}
fib.data <- wbm@meta.data[,c('Condition2','v6.labels','v7.labels','fibrosis.score')]

# head(fib.data)

fib.df2 <- data.frame(cluster = character(),
                      mean = numeric(),
                      sd = numeric())

cnt <- 1
for (i in unique(fib.data$v7.labels)){
  # print(i)
  
  temp.df <- fib.data[fib.data$v7.labels == i,]
  
  fib.df2[i,] <- c(i,mean(temp.df$fibrosis.score), sd(temp.df$fibrosis.score))
  
  cnt <- cnt + 1
}

fib.df2
```

### Looking at each large/generic cluster splitting by Experiment

#### Only whole bone marrow

Looking at just the whole bone marrow samples for Mpl Migr1

One-sided test to see if Mpl is higher

```{r}
fib.data$Exp <- ifelse(fib.data$Condition2 %in% c('Migr1','enrMigr1'), 'Migr1','Mpl')

fib.df2 <- data.frame(mean.Migr1 = numeric(),
                      mean.Mpl = numeric(),
                      t.stat = numeric(),
                      pval = numeric(),
                      BH.pval = numeric())

for (i in unique(fib.data$v7.labels)){
  # print(i)
  
  list1 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'Migr1',]$fibrosis.score
  list2 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'Mpl',]$fibrosis.score

  ttest <- t.test(list1,list2, alternative = 'less')
  
  fib.df2[i,] <- c(ttest$estimate[1],ttest$estimate[2],
               ttest$statistic, ttest$p.value, NA)
}

fib.df2$BH.pval <- p.adjust(fib.df2$pval,'BH')

fib.df2
```

#### Only whole bone marrow (Two-Sided)

Looking at just the whole bone marrow samples for Mpl Migr1

One-sided test to see if Mpl is higher

```{r}
fib.data$Exp <- ifelse(fib.data$Condition2 %in% c('Migr1','enrMigr1'), 'Migr1','Mpl')

fib.df2 <- data.frame(mean.Migr1 = numeric(),
                      mean.Mpl = numeric(),
                      t.stat = numeric(),
                      pval = numeric(),
                      BH.pval = numeric())

for (i in unique(fib.data$v7.labels)){
  # print(i)
  
  list1 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'Migr1',]$fibrosis.score
  list2 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'Mpl',]$fibrosis.score

  ttest <- t.test(list1,list2)
  
  fib.df2[i,] <- c(ttest$estimate[1],ttest$estimate[2],
               ttest$statistic, ttest$p.value, NA)
}

fib.df2$BH.pval <- p.adjust(fib.df2$pval,'BH')

temp.subset1 <- subset(wbm, Condition2 %in% c('Migr1','Mpl'))

VlnPlot(temp.subset1, 'fibrosis.score', pt.size = 0,
        group.by = 'v7.labels', split.by = 'Condition2')

fib.df2
```

#### Enriched and Non-enriched combined

One-sided test to see if Mpl score is greater

```{r}
fib.df2 <- data.frame(mean.Migr1 = numeric(),
                      mean.Mpl = numeric(),
                      t.stat = numeric(),
                      pval = numeric(),
                      BH.pval = numeric())

for (i in unique(fib.data$v7.labels)){
  # print(i)
  
  list1 <- fib.data[fib.data$v7.labels == i & fib.data$Exp == 'Migr1',]$fibrosis.score
  list2 <- fib.data[fib.data$v7.labels == i & fib.data$Exp == 'Mpl',]$fibrosis.score

  ttest <- t.test(list1,list2, alternative = 'less')
  
  fib.df2[i,] <- c(ttest$estimate[1],ttest$estimate[2],
               ttest$statistic, ttest$p.value, NA)
}

fib.df2$BH.pval <- p.adjust(fib.df2$pval,'BH')

fib.df2
```

### Looking at each large/generic cluster enriched vs non-enriched

#### Migr1

Two-sided test

```{r}
fib.df2 <- data.frame(mean.Migr1 = numeric(),
                      mean.enrMigr1 = numeric(),
                      t.stat = numeric(),
                      pval = numeric(),
                      BH.pval = numeric())

for (i in unique(fib.data$v7.labels)){
  # print(i)
  
  list1 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'Migr1',]$fibrosis.score
  list2 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'enrMigr1',]$fibrosis.score

  ttest <- t.test(list1,list2)
  
  fib.df2[i,] <- c(ttest$estimate[1],ttest$estimate[2],
               ttest$statistic, ttest$p.value, NA)
}

fib.df2$BH.pval <- p.adjust(fib.df2$pval,'BH')

fib.df2
```

Both Mono/Macro and MEP/ERP show up as significant

#### Mpl

Two-sided test

```{r}
fib.df2 <- data.frame(mean.Mpl = numeric(),
                      mean.enrMipl = numeric(),
                      t.stat = numeric(),
                      pval = numeric(),
                      BH.pval = numeric())

labels.with.enough.cells <- unique(fib.data$v7.labels)[-c(2,8)]

for (i in labels.with.enough.cells){
  # print(i)
  
  list1 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'Mpl',]$fibrosis.score
  list2 <- fib.data[fib.data$v7.labels == i & fib.data$Condition2 == 'enrMpl',]$fibrosis.score

  ttest <- t.test(list1,list2)
  
  fib.df2[i,] <- c(ttest$estimate[1],ttest$estimate[2],
               ttest$statistic, ttest$p.value, NA)
}

fib.df2$BH.pval <- p.adjust(fib.df2$pval,'BH')

fib.df2
```

Granulocyte, Mono/Macro, Erythroid and MEP/ERP all have different distrubtions of fibrosis scores between Mpl and enrMpl

## Second Gene Set

From Supp Table 5 from the Oxford (paper)[https://www.cell.com/molecular-cell/fulltext/S1097-2765(20)30234-3?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS1097276520302343%3Fshowall%3Dtrue#supplementaryMaterial]


```{r}
fib.genes2 <- c('Tgfb1','Il12a','Il15','Il15','Il1b','Actr5','Rab37','C20orf194',
                'Tollip','Gosr2')

fib.genes2 %in% rownames(wbm)

# rownames(wbm)[grep('C20', rownames(wbm), ignore.case = T)]

fib.genes2 <- fib.genes2[fib.genes2 %in% rownames(wbm)]
```


```{r}

cell.total.score <- colSums(cnts)

cnts <- cnts[rownames(cnts) %in% fib.genes2,]

cell.raw.score <- colSums(cnts)

fib.df <- data.frame(cell = colnames(cnts),
                     cell.total.score = cell.total.score,
                     fib.total.score = cell.raw.score)

fib.df$weighted.score <- fib.df$fib.total.score / fib.df$cell.total.score

ggplot(fib.df, aes (x = weighted.score)) + geom_density()

fib.df$scaled.score <- (fib.df$weighted.score - mean(fib.df$weighted.score)) /
  sd(fib.df$weighted.score)

wbm$fibrosis.score <- fib.df$scaled.score
wbm$fibrosis.score.unstandardized <- fib.df$weighted.score

FeaturePlot(wbm, features = 'fibrosis.score', cols = c('lightgrey','red'))
FeaturePlot(wbm, features = 'fibrosis.score.unstandardized')
```

## Tyapas Lab List

```{r}
l1 <- read.csv('./data/tapas_lab/rs.CD34.MF_v_others_edgeR_allGenes (1).csv', 
               header = T)

head(l1)

cd34 <- l1[l1$FDR < 0.05,]

cd34.up.genes <- cd34[cd34$logFC > 0,]$gene.symbol
cd34.down.genes <- cd34[cd34$logFC < 0,]$gene.symbol

pbmc <- read.csv('./data/tapas_lab/rs.PBMC.BMMC.MF_v_others_edgeR_allGenes (1).csv',
                 header = T)
pbmc <- pbmc[pbmc$FDR < 0.05,]
pbmc.up.genes <- pbmc[pbmc$logFC > 0,]$gene.symbol
pbmc.down.genes <- pbmc[pbmc$logFC < 0,]$gene.symbol

de.genes <- read.xlsx('./data/v3/within.cluster.DEGs.Migr1.vs.Mpl.LARGE.GENERIC.CLUSTERS.xlsx',
                      sheetName = 'All-Clusters')
table(de.genes$cluster)


mep.de.genes <- de.genes[de.genes$cluster == 'MEP/MCP',]$gene

summary(toupper(mep.de.genes) %in% pbmc$gene.symbol)
summary(toupper(mep.de.genes) %in% cd34$gene.symbol)
```




```{r}
Idents(wbm) <- wbm$v7.labels
VlnPlot(wbm, 'Itga2b')
```

# Trajectory

```{r}

```

