---
title: "v5.CanIntegrateFromStart.."
author: "D. Ford Hannum Jr."
date: "12/11/2020"
output: 
        html_document:
                toc: true
                toc_depth: 3
                number_sections: false
                theme: united
                highlight: tango
---

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(Seurat)
library(ggplot2)
library(data.table)
library(MAST)
library(SingleR)
library(dplyr)
library(tidyr)
library(limma)
library(scRNAseq)
library(ggpubr)
```

From talking with Jun it may not be a bad idea to integrate from the beginning. Going to look at cross tabulation between the original experiments (which I started again in v3.WriteUp.Rmd), the Nbeal aged-control (which I'll generate in this document), and the integrated data from v2.1.Integration.UMAP.html.

Also going to look at how the cluster centroids look across PC space.

```{r loading data}
migr.mpl <- readRDS('./data/v3/mpl.migr1.new.object.rds')
int <- readRDS('./data/v2/lesser.combined.integrated.NAMED.rds')

int$idents2 <- ifelse(int$idents != 'HSPC' & int$idents != 'CMP' & int$idents != 'MEP/Mast', as.character(int$idents),
                      ifelse(int$idents == 'MEP/Mast', 'MEP/MCP',
                             ifelse(int$idents == 'HSPC','Multipotent Progenitor', 'Progenitor')))

Idents(int) <- int$idents2

int$cluster <- int$new_cluster_IDs

int$cluster <- as.character(int$cluster)

int$cluster <- ifelse(int$cluster == '?GMP', 'Gran-4',
                      ifelse(int$cluster == '?CMP/Neutro', 'Progenitor', int$cluster))

int@meta.data[int$cluster == 'Megakaryocyte',]$cluster <- int@meta.data[int$cluster == 'Megakaryocyte',]$idents2

table(int$cluster, Idents(int))

## 

migr.mpl$immgen.labels <- factor(migr.mpl$seurat_clusters,
                                 labels = c('Gran-1','Gran-2','B cell-1','Basophil','Stem Cell','B cell-2',
                                            'Monocyte','Gran-3','Macrophage','B cell-3','Stem Cell','NA','T cell',
                                            'B cell-4','NA','DC'))

migr.mpl$labels <- factor(migr.mpl$seurat_clusters,
                                 labels = c('Gran-1','Gran-2','B cell-1','MEP/Mast','Prog','?',
                                            'Monocyte','Gran-3','Macrophage','B cell-3','Stem Cell','Erythrocyte','T cell',
                                            'Erythrocyte','MK','B cell-4'))

DimPlot(migr.mpl, reduction = 'umap', group.by = 'labels', label = T, repel = T)
```

# Looking at the different centroids for migr1 mpl

Focusing just on the migr1 mpl data points

1. cluster centroids from the pre-integrated data
2. cluster centroids for the post-integrated data (RNA)

# Full Clusters

Not making any distinction between either Migr1 vs Mpl or Enriched vs Non-enriched

## Cross Tabulation

```{r ct}
head(migr.mpl@meta.data)

# Will maybe use later on
migr.mpl$exp <- ifelse(migr.mpl$Condition %in% c('enrMigr1','Migr1'), 'Migr1','Mpl')

head(int@meta.data)

summary(as.factor(int$orig.ident))

DimPlot(int, split.by = 'orig.ident')

DimPlot(int, group.by = 'orig.ident', cols = c('red','grey'))
DimPlot(int, group.by = 'orig.ident', cols = c('grey','red'))

int.migr.mpl <- subset(int, orig.ident == 'Mpl')

Idents(migr.mpl) <- migr.mpl$labels

summary(rownames(migr.mpl@meta.data) == tstrsplit(rownames(int.migr.mpl@meta.data),'_',keep = 2)[[1]])

ct2 <- data.frame(orig.cluster = migr.mpl$seurat_clusters,
                 orig.cluster.label = migr.mpl$labels,
                 int.cluster = int.migr.mpl$seurat_clusters,
                 int.cluster.label = int.migr.mpl$cluster)

head(ct2)

ct <- as.data.frame(table(ct2$orig.cluster.label, ct2$int.cluster.label))

head(ct)

colnames(ct) <- c('Orig.Cluster','Int.Cluster','Freq')

head(ct)

ct$Orig.Cluster.total <- NA
for (i in unique(ct$Orig.Cluster)){
        # print(i)
        
        ct[ct$Orig.Cluster ==i,]$Orig.Cluster.total <- sum(ct[ct$Orig.Cluster ==i,]$Freq)
}

ct$Int.Cluster.total <- NA

for (i in unique(ct$Int.Cluster)){
        ct[ct$Int.Cluster == i,]$Int.Cluster.total <- sum(ct[ct$Int.Cluster ==i,]$Freq)
}

ct$Orig.Cluster.perc <- ct$Freq/ct$Orig.Cluster.total
ct$Int.Cluster.perc <- ct$Freq/ct$Int.Cluster.total

int.levels <- c('Gran-1','Gran-2','Gran-3','Gran-4','Progenitor',
                'Monocyte','Macrophage','Erythrocyte','Megakarycoyte',
                'MEP/Mast','B cell-1','B cell-2', 'T cell',
                'B cell-3','B cell-4', 'Multipotent Progenitor')

ct$Int.Cluster2 <- factor(ct$Int.Cluster, levels = int.levels)

orig.levels <- c('Gran-2','Gran-3','Gran-1','?','Prog','Stem Cell','Monocyte',
                 'B cell-4','Macrophage','MK','Erythrocyte','MEP/Mast',
                 'B cell-1','B cell-3','T cell')

ct$Orig.Cluster2 <- factor(ct$Orig.Cluster, levels = orig.levels)

ggplot(ct, aes(x = Orig.Cluster2, y = Int.Cluster2, fill = Orig.Cluster.perc)) +
        geom_tile() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = .5) +
        geom_text(aes(label = Freq))

table(ct$Int.Cluster, ct$Int.Cluster2)
```

#### Aside looking at MK cluster

This is something that I've mean to look at but forgot until I saw the above figure

Oddly we see the integrated cluster of Macrophages also contains all the original MKs (or just about). When looking at the expression of MK markers we see that a tail of the Macrophage cluster looks like it belongs to the corresponding mk cluster.

```{r looking at the oddity of hte MK cluster}

DimPlot(int, label = T, repel = T, group.by = 'cluster')

int.migr.mpl$orig.label <- migr.mpl$labels

temp <- subset(int, cluster %in% c('Megakarycoyte','Macrophage'))
DimPlot(temp)

FeaturePlot(temp, features = 'Vwf', split.by = 'cluster', pt.size = .11)

DimPlot(int.migr.mpl, label = T, repel = T, group.by = 'cluster')


# Looking at just hte migr1.mpl cells

temp <- subset(int.migr.mpl, cluster %in% c('Megakarycoyte','Macrophage'))

DimPlot(temp)

FeaturePlot(temp, features = 'Vwf', split.by = 'cluster', pt.size = .11)

FeaturePlot(temp, features = 'Vwf', split.by = 'orig.label', pt.size = .11)

temp <- subset(temp, orig.label %in% c('Macrophage','MK'))

FeaturePlot(temp, features = 'Vwf', split.by = 'orig.label', pt.size = )

DimPlot(temp, label= T, repel = T) + NoLegend()
```

Also interestingly there seems to be many macrophage cells occuppying the same space as our granulocyte clusters...should we be excluding the granulocytes (1-3), b cells and t cells to refine our analysis?